---
title: "0-ModelFormulations"
author: "Galina M. Jönsson"
date: "17/06/2021"
output:   
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### For each of 3 models: 
1. **Model A**: categorical List length   
2. **Model B**: mixed list length   
3. **Model C**: mixed list length without year effect for NHCs    

# Model fomulation and justification 

## **Model A**   
**Model A** specifies that list length should be considered as a categorical variable. Per default, [sparta] (https://github.com/BiologicalRecordsCentre/sparta) has three list length classes:  lists of length 1, 2-3, and 4 and over. I have manually changes the list lengths of all BNM, UKBMS and NHC data to be treated as three separate list lengths. This model formulation is equivalent to our [previous paper] (https://doi.org/10.1111/icad.12494). However, here we 





## **Model B**   
**Model B** list length as both continuous and categorical variable. 



## **Model C**   
**Model C** list length as both continuous and categorical variables with varying year effects between list lengths


```{r load-data}
ModelA_A_urticae<- readRDS("../outputs/catLL-outputs/results_Aglais_urticae_watson_catLL.rds")
ModelB_A_urticae <- readRDS("../outputs/mixLL-outputs/results_Aglais_urticae_crick_mixLL.rds")
ModelC_A_urticae <- readRDS("../outputs/mixLL2-outputs/results_Aglais_urticae_crick_mixLL2.rds")
```

# JAGS observation model code


#### Model A
```{r jags-DetProb-code-A, eval=FALSE, echo=TRUE}
for(j in 1:nvisit) {    
  y[j] ~ dbern(Py[j])    
  Py[j]<- z[Site[j],Year[j]]*p[j]    
  logit(p[j]) <-  alpha.p[Year[j]] + dtype2.p * DATATYPE2[j] + dtype3.p * DATATYPE3[j]    
} }   
Fully observed variables:   
 DATATYPE2 DATATYPE3 Site Year nsite nvisit nyear y  
```


#### Model B
```{r jags-DetProb-code-B, eval=FALSE, echo=TRUE}
for(j in 1:nvisit) {   
  y[j] ~ dbern(Py[j])   
  Py[j]<- z[Site[j],Year[j]] * p[j]   
  logit(p[j]) <-  alpha.p[Year[j]] + LL.p * logL[j] + dtype2.p * DATATYPE2[j] + dtype3.p*DATATYPE3[j]   
}   
Fully observed variables:   
 DATATYPE2 DATATYPE3 Site Year dtype2p_max dtype2p_min logL nsite nvisit nyear y    
```

#### Model C
```{r jags-DetProb-code-C, eval=FALSE, echo=TRUE}
for(j in 1:nvisit) {
  y[j] ~ dbern(Py[j])
  Py[j]<- z[Site[j],Year[j]] * p[j]
  logit(p[j]) <-  alpha.p[Year[j]] * (1-DATATYPE3[j]) + LL.p * logL[j] + dtype2.p * DATATYPE2[j] + dtype3.p * DATATYPE3[j]
} }
Fully observed variables:
 DATATYPE2 DATATYPE3 Site Year dtype2p_max dtype2p_min logL nsite nvisit nyear y 
```




# Detection probability

```{r jags-DetProb-code, echo=FALSE}
require(ggplot2)
require(sparta)


plot_DetectionOverTime(ModelA_A_urticae,
                              min.yr = 1900, 
                                    legend_labels = c("BNM/Year_Effect", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle(label = (expression(paste(italic("Aglais urticae"), " - Model A"))), 
          subtitle = paste("Small Tortoiseshell")) + 
  theme(plot.title=element_text(face="italic")) +
  theme(text=element_text(size=12)) 



plot_DetectionOverTime(ModelB_A_urticae, 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle(label = (expression(paste(italic("Aglais urticae"), " - Model B"))), 
          subtitle = paste("Small Tortoiseshell")) + 
  theme(plot.title=element_text(face="italic")) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(ModelC_A_urticae, 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type",
                       mixLL2 = TRUE) +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle(label = (expression(paste(italic("Aglais urticae"), " - Model C"))), 
          subtitle = paste("Small Tortoiseshell")) + 
  theme(plot.title=element_text(face="italic")) +
  theme(text=element_text(size=12))

```





