---
title: "03-OccupancyTrends_new"
author: "Galina M. Jönsson"
date: "16/11/2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---
## Set up

#### Load required packages
```{r load-packages, cache=TRUE, message=FALSE}
# Packages for modeling
library(devtools)
list.of.packages <- c("minqa", "lme4", "gtools", "gtable", "scales",
                      "assertthat", "magrittr", "tibble", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
#install_github(repo = 'biologicalrecordscentre/BRCindicators')
#install_github('biologicalrecordscentre/sparta')
library("sparta")
library("BRCindicators")
```




# Inspect output

## Convergance
I run the analyses on a computer cluster and must fist test whether all annual estimates have converged. 

```{r test-Rhat, cache=TRUE, message=FALSE}
# Source the function that loops through all model outputs and checks convergence, i.e. is Rhat >1.1?
source("./function-summarise_Rhat.R")

# Run function
RhatSummary <- summarise_Rhat("../outputs/mixLL-outputs/", verbose=FALSE)
# Find the species which need more iterations (converged-column == "N")
require(dplyr)
NotConvergedSpp <- subset(RhatSummary, grepl("N", converged))
NotConvergedSpp
```
*Aglais io*, *Apatura iris*, *Ochlodes sylvanus* and *Pyronia tithonus* did not converge when modeled with 50,000 iterations. All were subsequentually modeled with 80,000 iterations and as seen avove, only *Pyronia tithonus* still did not converge. The mixlistlength model outputs for *Pyronia tithonus* will not be considered further here. 



## Visualise occupancy

```{r plot-Occupancy, cache=TRUE, message=FALSE}
require(sparta) 

# Create empty list
list <- list() 
names <- list()

# Input directory
input_dir <- "../outputs/mixLL-outputs_Converged/"   # Only converged models used
files <- list.files(path = paste(input_dir), ignore.case = TRUE, pattern = '\\.rds$') # list of the files to loop through

# Loop through the files and list occupancy plots and species names in two separate lists
for(i in 1:length(files)){
  # read the rds file
  out <- readRDS(file.path(input_dir, files[i])) 
  list[[i]] <- plot(out)
  names[[i]] <- out$SPP_NAME
}


list
``` 



## Visualise detection

### Habitat specialists
```{r plot-Detection_HS, cache=TRUE, message=FALSE, warning=FALSE}
# Create empty list
list2 <- list() 

# Input directory
input_dir <- "../outputs/mixLL-outputs_HS/"
files2 <- list.files(path = paste(input_dir), ignore.case = TRUE, pattern = '\\.rds$') # list of the files to loop through

for(i in 1:length(files2)){
  # read the rds file
  out <- readRDS(file.path(input_dir, files2[i])) 
  list2[[i]] <- plot_DetectionOverTime(out, out$SPP_NAME, min.yr = 1900, legend_labels = c("BNM_LL1/Year_Effect", "UKBMS", "NHCs", "BNM_LL5"), legend_title = "Data Type")
}

list2
```


### Wider Countryside Species
```{r plot-Detection_WCS, cache=TRUE, message=FALSE, warning=FALSE}
# Create empty list
list3 <- list() 

# Input directory
input_dir <- "../outputs/mixLL-outputs_WCS/"
files3 <- list.files(path = paste(input_dir), ignore.case = TRUE, pattern = '\\.rds$') # list of the files to loop through

for(i in 1:length(files3)){
  # read the rds file
  out <- readRDS(file.path(input_dir, files3[i])) 
  list3[[i]] <- plot_DetectionOverTime(out, out$SPP_NAME, min.yr = 1900, legend_labels = c("BNM_LL1/Year_Effect", "UKBMS", "NHCs", "BNM_LL5"), legend_title = "Data Type")
}

list3
```
\n
\n
\n


### Select species for further model comparison
\n
I select 15 species to run using the mixlistlength2 observation sub-model formulation based on the following criteria and methodology: 
\n

1. Geographic coverage: subset all species into three categories: low (<0.25 percentile), medium (0.25-0.75 percentile) and high (>0.75 percentile) occupancy today   
2. Habitat generalists (WCS) or specialist (HS): group each geographic coverage category into either specialists or generalists, resulting in six groups       
3. Go through the species in each of the six groups and select species based on whether they show much interannual variation in detection probability. Species that show much variation and uncertainty are expected to benefit most from an improved detection sub-model formulation       
\n
```{r MixLL2-select1, cache=TRUE, message=FALSE, warning = FALSE, eval=TRUE}
#### 1 

# Source the function that loops through all model outputs and summarises
source("./function-sumOccDet_output.R")

# Run function
output <- sumOccDet_output("../outputs/mixLL-outputs_Converged/", categoriseOccu = TRUE, quantileProbs = c(0, 0.25, 0.75, 1))




#### 2

# Load generalists vs. specialist look-up table
CatLookUp <- read.csv("../data/auxiliaryData/MSI_lookup.csv")

# Substitute blank space in species name for punctuation
CatLookUp$Species <- gsub(" ", ".", CatLookUp$Species)
CatLookUp$HS_WCS <- as.factor(CatLookUp$HS_WCS)

# Merge with output
output <- merge(output, CatLookUp, by.x = "species_name", by.y = "Species", all.x = TRUE, all.y = FALSE)
```


\n
\n
\n
\n



**Low geographic coverage + HS** \n
Please note that there are no species in the low geographic coverage category that are habitat generalists. Hence, I will select four instead of two species here.   
```{r MixLL2-select-Low-HS, cache=TRUE, message=FALSE, eval=FALSE}
(subset((subset(output, category=="1")), HS_WCS=="HS"))
```
- I will chose the three species with the largest mean annual detection probability standard deviations (*Erebia epiphron*, *Carterocephalus palaemon* and *Melitaea athalia*) and the sixth (of 13) most variable species, *Polyommatus bellargus*, which has most records of all species in this group.   

\n
\n
\n


**Medium geographic coverage + HS**
```{r MixLL2-select-Medium-HS, cache=TRUE, message=FALSE, eval=FALSE}
(subset((subset(output, category=="2")), HS_WCS=="HS"))
```
- I will chose the species with the largest mean annual detection probability standard deviation (*Thecla betulae*) and fourth (of 16) most variable species, *Limenitis camilla*, as it has among the most records of the species in this group and in contrast to all other species has shown recent range expansions. 

\n
\n
\n

**Medium geographic coverage + WCS**
```{r MixLL2-select-Medium-WCS, cache=TRUE, message=FALSE, eval=FALSE}
(subset((subset(output, category=="2")), HS_WCS=="WCS"))
```
- I will chose the species with the largest mean annual detection probability standard deviation (*Polygonia c-album*) and the third (of 10) most variable species, *Erebia aethiops*, as it has among the fewest records of the species in this group.

\n
\n
\n


**High geographic coverage + WCS**
Please note that there are no species in the high geographic coverage category that are habitat specialists. Hence, I will select four instead of two species here. 
```{r MixLL2-select-High-WCS, cache=TRUE, message=FALSE, eval=FALSE}
(subset((subset(output, category=="3")), HS_WCS=="WCS"))
```
- I will chose the three species with the largest mean annual detection probability standard deviations (*Aglais urticae*, *Pararge aegeria* and *Lycaena phlaeas*) and the fourth (of 13) most variable species, *Pieris rapae*, as the genus of the third (*Aglais io*) is already represented in the sample.

\n
\n
\n



**In short, the species I will run using the MixLL2 formulation are:**   
1. *Erebia epiphron*   
2. *Carterocephalus palaemon*   
3. *Melitaea athalia*   
4. *Polyommatus bellargus*   
5. *Thecla betulae*    
6. *Limenitis camilla*   
7. *Polygonia c-album*   
8. *Erebia aethiops*   
9. *Aglais urticae*   
10. *Pararge aegeria*   
11. *Lycaena phlaeas*   
12. *Pieris rapae*   


\n
\n
\n
\n
\n
\n
\n
\n
\n
\n


# Multispecies indicators

## BMA
```{r BMA, cache=TRUE, message=FALSE, eval=FALSE}
# Load my arrayBuilder function
source("./function-arrayBuilder.R")


############## All 52 species
# Use function to build a three-dimensional matrix. The dimensions of this array represent species, years, and iterations.
lambdaArray_all <- arrayBuilder(input_dir = "../outputs/mixLL-outputs/", nsp=54, startyr=1, stopyr=110, iter=9999)

# Ensure values are bounded by 0 and 1
lambdaArray_all[lambdaArray_all == 1] <- 0.99999999999999999
lambdaArray_all[lambdaArray_all == 0] <- 0.00000000000000001

spp <- names(lambdaArray_all[,1,1])
trends_BMA_all <- data.frame(species = as.factor(rep(names(lambdaArray_all[,1,1]), each = 110)),
                             year = as.numeric(rep(1:110, 54)),
                             index = as.numeric(NA),
                             se = as.numeric(NA))

# Loop through an elog values (required by indicator function)
count <- 0
for(i in 1:length(spp)){

  for(j in 1:110){
      count <- count + 1
      trends_BMA_all[count,"index"] <- mean(log((lambdaArray_all[i,j,]), base = exp(1)))
      trends_BMA_all[count,"se"] <- sd(log((lambdaArray_all[i,j,]), base = exp(1)))/sqrt(length((lambdaArray_all[i,j,])))
  }
}

# Run indicator function
bma_indicator_all2k <- bma(trends_BMA_all,
                         model = "smooth_det2",
                         parallel = TRUE,
                     incl.model = TRUE,
                     n.iter = 20000, 
                     m.scale = 'loge',
                     num.knots = 50,
                     rescaleYr = 76, 
                     n.thin = 5)



############## Habitat specialists (29 species)

# Use function to build a three-dimensional matrix. The dimensions of this array represent species, years, and iterations.
lambdaArray_HS <- arrayBuilder(input_dir = "../outputs/CtagWatsonCrick_results_HS/", nsp=29, startyr=1, stopyr=110, iter=9999)

# Ensure values are bounded by 0 and 1
lambdaArray_WCS[lambdaArray_WCS == 1] <- 0.99999999999999999
lambdaArray_WCS[lambdaArray_WCS == 0] <- 0.00000000000000001


trends_BMA_WCS <- data.frame(species = as.factor(rep(names(lambdaArray_WCS[,1,1]), each = 110)),
                             year = as.numeric(rep(1:110, 23)),
                             index = as.numeric(NA),
                             se = as.numeric(NA))

# Loop through an elog values (required by indicator function)
spp <- names(lambdaArray_WCS[,1,1])
count <- 0
for(i in 1:length(spp)){

  for(j in 1:110){
      count <- count + 1
      trends_BMA_WCS[count,"index"] <- mean(log((lambdaArray_WCS[i,j,]), base = exp(1)))
      trends_BMA_WCS[count,"se"] <- sd(log((lambdaArray_WCS[i,j,]), base = exp(1)))/sqrt(length((lambdaArray_WCS[i,j,])))
  }
}


# Run indicator function
bma_indicator_1yrHS <- bma(trends_BMA_HS,
                     parallel = TRUE,
                     model = "smooth_det",
                     n.iter = 20000, 
                     num.knots = 50, 
                     rescaleYr = 1,
                     n.thin = 5,
                     m.scale = 'loge')





############## Habitat generalists (23 species)

# Use function to build a three-dimensional matrix. The dimensions of this array represent species, years, and iterations.
lambdaArray_WCS <- arrayBuilder(input_dir = "../outputs/CtagWatsonCrick_results_WCS/", nsp=23, startyr=1, stopyr=110, iter=9999)

# Ensure values are bounded by 0 and 1
lambdaArray_HS[lambdaArray_HS == 1] <- 0.99999999999999999
lambdaArray_HS[lambdaArray_HS == 0] <- 0.00000000000000001

spp <- names(lambdaArray_HS[,1,1])
trends_BMA_HS <- data.frame(species = as.factor(rep(names(lambdaArray_HS[,1,1]), each = 110)),
                             year = as.numeric(rep(1:110, 29)),
                             index = as.numeric(NA),
                             se = as.numeric(NA))

# Loop through an elog values (required by indicator function)
count <- 0
for(i in 1:length(spp)){

  for(j in 1:110){
      count <- count + 1
      trends_BMA_HS[count,"index"] <- mean(log((lambdaArray_HS[i,j,]), base = exp(1)))
      trends_BMA_HS[count,"se"] <- sd(log((lambdaArray_HS[i,j,]), base = exp(1)))/sqrt(length((lambdaArray_HS[i,j,])))
  }
}

# Run indicator function
bma_indicator_1yrWCS <- bma(trends_BMA_WCS,
                     parallel = TRUE,
                     model = "smooth_det",
                     n.iter = 20000, 
                     num.knots = 50, 
                     rescaleYr = 1,
                     n.thin = 5,
                     m.scale = 'loge')



####### Plot


plot_indicator(indicator = bma_indicator_all2k[,'Index'],
               CIs = bma_indicator_all2k[,c(3,4)])

plot_indicator(indicator = bma_indicator_1yrHS[,'Index'],
               CIs = bma_indicator_1yrHS[,c(3,4)])

plot_indicator(indicator = bma_indicator_1yrWCS[,'Index'],
               CIs = bma_indicator_1yrWCS[,c(3,4)])

source("./fig-plot_LMSI.R")
require(ggplot2)
plot_LMSIgal(indicator1 = (indicator = bma_indicator_all2k[,'Index']),
               CIs1 = (bma_indicator_all2k[,c(3,4)]),
             indicator2 = (indicator = bma_indicator_1yrHS[,'Index']),
               CIs2 = (bma_indicator_1yrHS[,c(3,4)]),
             indicator3 = (indicator = bma_indicator_1yrWCS[,'Index']),
               CIs3 = (bma_indicator_1yrWCS[,c(3,4)]))

```



## Monte Carlo Multi-Species Indicator
Statistics Neatherlands' method to account for sampling error of species indices in the calculation of multi-species indicators based on Monte Carlo simulation of annual species indices (Soldaat et al., 2017; https://doi.org/10.1016/j.ecolind.2017.05.033).

### Monte Carlo MSI: Habitat specialists vs. Wider countryside species
```{r MSI, cache=TRUE, message=FALSE, eval=FALSE}
# Load summarise_Rhat function
source("./function-summarise_occDetRds_BMA.R")


# Run function to get data in appropriate format (53 species)
MSI_data <- summarise_occDetRdsBMA("../outputs/mixLL-outputs/")
# Year starts with 1 to 110 so add 1899
MSI_data$year <- MSI_data$year + 1899


############# Specialist vs. generalist species 
CatLookUp <- read.csv("../data/auxiliaryData/MSI_lookup.csv")
CatLookUp$Species <- gsub(" ", ".", CatLookUp$Species)
CatLookUp$HS_WCS <- as.factor(CatLookUp$HS_WCS)

# Habitat specialists (HS) # 29 species
require(dplyr)
names.use <- unique(MSI_data$species[MSI_data$species %in% ((subset(CatLookUp, HS_WCS == "HS"))$Species)])
#trends_summary_HS <- trends_summary[, (append("year", names.use))]
MSI_data_HS <- MSI_data[MSI_data$species %in% names.use,]

# Wider countryside species (WCS) # 24 species
names.use <- unique(MSI_data$species[MSI_data$species %in% ((subset(CatLookUp, HS_WCS == "WCS"))$Species)])
#trends_summary_WCS <- trends_summary[, (append("year", names.use))]
MSI_data_WCS <- MSI_data[MSI_data$species %in% names.use,]


# Species index values need to be 100 in the base year. Here I use
# the first year as my base year and rescale to 100. The standard error
# in the base year should be 0.
min_year <- min(MSI_data$year)

for(sp in unique(MSI_data$species)){
  subset_data <- MSI_data[MSI_data$species == sp, ]
  multi_factor <- 100 / subset_data$index[subset_data$year == min_year]
  MSI_data$index[MSI_data$species == sp] <- MSI_data$index[MSI_data$species == sp] * multi_factor
  MSI_data$se[MSI_data$species == sp] <- MSI_data$se[MSI_data$species == sp] * multi_factor
  MSI_data$se[MSI_data$species == sp][1] <- 0
}

###### HS
min_year <- min(MSI_data_HS$year)

for(sp in unique(MSI_data_HS$species)){
  subset_data <- MSI_data_HS[MSI_data_HS$species == sp, ]
  multi_factor <- 100 / subset_data$index[subset_data$year == min_year]
  MSI_data_HS$index[MSI_data_HS$species == sp] <- MSI_data_HS$index[MSI_data_HS$species == sp] * multi_factor
  MSI_data_HS$se[MSI_data_HS$species == sp] <- MSI_data_HS$se[MSI_data_HS$species == sp] * multi_factor
  MSI_data_HS$se[MSI_data_HS$species == sp][1] <- 0
}

###### WCS
min_year <- min(MSI_data_WCS$year)

for(sp in unique(MSI_data_WCS$species)){
  subset_data <- MSI_data_WCS[MSI_data_WCS$species == sp, ]
  multi_factor <- 100 / subset_data$index[subset_data$year == min_year]
  MSI_data_WCS$index[MSI_data_WCS$species == sp] <- MSI_data_WCS$index[MSI_data_WCS$species == sp] * multi_factor
  MSI_data_WCS$se[MSI_data_WCS$species == sp] <- MSI_data_WCS$se[MSI_data_WCS$species == sp] * multi_factor
  MSI_data_WCS$se[MSI_data_WCS$species == sp][1] <- 0
}
# Our first year is now indexed at 100


###### ALL species
min_year <- min(MSI_data$year)

for(sp in unique(MSI_data$species)){
  subset_data <- MSI_data[MSI_data$species == sp, ]
  multi_factor <- 100 / subset_data$index[subset_data$year == min_year]
  MSI_data$index[MSI_data$species == sp] <- MSI_data$index[MSI_data$species == sp] * multi_factor
  MSI_data$se[MSI_data$species == sp] <- MSI_data$se[MSI_data$species == sp] * multi_factor
  MSI_data$se[MSI_data$species == sp][1] <- 0
}
# Our first year is now indexed at 100



#msi_out <- msi(data,
#               nsim = 500, # The number of Mote Carlo simulations
#               SEbaseyear = 1950, # The year to index on
#               plotbaseyear = 1970, # The year to set as 100 in plots
#               index_smoot = 'INDEX', # plotbaseyear uses MSI not trend
#               span = 0.7, # 'wigglyness' of line, between 0 and 1
#               lastyears = 1910, # last X years of time series for short-term trends
#               maxCV = 10, # maximum allowed Coefficient of Variation 
#               changepoint = 1970, # compare trends before and after this year
#               truncfac = 8, # max year-to-year index ratio
#               TRUNC = 5, #set all indices below TRUNC to this
#               plot = FALSE # should the plots be returned?)
#               )

require(BRCindicators)
msi_out_all <- msi(MSI_data,
                   nsim = 10000, # The number of Mote Carlo simulations
                   changepoint = 1976, # compare trends before and after this year
                   index_smooth = 'INDEX', # plotbaseyear uses MSI not trend
                   lastyears = 34,
                   span = 0.5,
                   plot = FALSE # should the plots be returned?)
                   )

### HS
msi_out_HS <- msi(MSI_data_HS,
                   nsim = 10000, # The number of Mote Carlo simulations
                   changepoint = 1976, # compare trends before and after this year
                   index_smooth = 'INDEX', # plotbaseyear uses MSI not trend
                   lastyears = 34,
                   span = 0.5,
                   plot = FALSE # should the plots be returned?)
                   )

### WSC
msi_out_WCS <- msi(MSI_data_WCS,
                   nsim = 10000, # The number of Mote Carlo simulations
                   changepoint = 1976, # compare trends before and after this year
                   index_smooth = 'INDEX', # plotbaseyear uses MSI not trend
                   lastyears = 34,
                   span = 0.5,
                   plot = FALSE # should the plots be returned?)
                   )


# Source the plotting function
source("./fig-plot_MSI.R")

plot.MSIgal(msi_out_all, msi_out_HS, msi_out_WCS)
plot.MSIgalOne(msi_out_all, msi_out_HS, msi_out_WCS)




msi_out_all$results$Trend<- msi_out_all$results$Trend+4.53
msi_out_all$results$lower_CL_trend<-msi_out_all$results$lower_CL_trend+4.53
msi_out_all$results$upper_CL_trend<- msi_out_all$results$upper_CL_trend+4.53
plot.MSIgal_Linnean(msi_out_all)

msi_out_HS$results$Trend<- msi_out_HS$results$Trend-9.15 
msi_out_HS$results$lower_CL_trend<-msi_out_HS$results$lower_CL_trend-9.15 
msi_out_HS$results$upper_CL_trend<- msi_out_HS$results$upper_CL_trend-9.15 
plot.MSIgal2_Linnean(msi_out_HS, msi_out_WCS)
```


## Lamda Multi-Species Indicator
From DEFRA/CEH. Calculates an indicator using growth rates from one year to the next. Advantages over the conventional approach to constructing indicators is that the categorisation of species as ‘increasing’ or ‘decreasing’ can be made from the same set of data (the growth rates) as the construction of the indicator. ALSO uses the actual iterations for each species specific estimate as opposed to mean+se. 

```{r lambdaIndicator, cache=TRUE, message=FALSE, eval=FALSE}
# Load my arrayBuilder function
source("./function-arrayBuilder.R")

# Use function to build a three-dimensional matrix. The dimensions of this array represent species, years, and iterations.
lambdaArray_all <- arrayBuilder(input_dir = "../outputs/mixLL-outputs/", nsp=53, startyr=1, stopyr=110, iter=9999)

lambdaArray_HS <- arrayBuilder(input_dir = "../outputs/mixLL-outputs_HS/", nsp=29, startyr=1, stopyr=110, iter=9999)

lambdaArray_WCS <- arrayBuilder(input_dir = "../outputs/mixLL-outputs_WCS/", nsp=24, startyr=1, stopyr=110, iter=9999)

# Ensure values are bounded by 0 and 1
# summary(lambdaArray_all > 1)
# summary(lambdaArray_all < 0)

# Run the lambda_interpolation method on this data
lambdaIndicator_all <- lambda_indicator(lambdaArray_all)

# Run the lambda_interpolation method on this data
lambdaIndicator_HS <- lambda_indicator(lambdaArray_HS)

# Run the lambda_interpolation method on this data
lambdaIndicator_WCS <- lambda_indicator(lambdaArray_WCS)

# Plot the indicator
source("./fig-plot_LMSI.R")
plot_LMSIgal(indicator1 = (lambdaIndicator_all$summary[,'indicator']),
               CIs1 = (lambdaIndicator_all$summary[,c('lower' ,'upper')]),
             indicator2 = (lambdaIndicator_HS$summary[,'indicator']),
               CIs2 = (lambdaIndicator_HS$summary[,c('lower' ,'upper')]),
             indicator3 = (lambdaIndicator_WCS$summary[,'indicator']),
               CIs3 = (lambdaIndicator_WCS$summary[,c('lower' ,'upper')]))
```


## Growth rate between time periods

```{r percentdif, cache=TRUE, message=FALSE, eval=FALSE}
# Load my function summarising percent differences between all successive years and giving the average 
source("./function-summarise_percentdif.R")
out <- summarise_percentdif("../outputs/mixLL-outputs/", cutoff = 77)

# Use function to build a three-dimensional matrix. The dimensions of this array represent species, years, and iterations.
lambdaArray_all_1TP <- arrayBuilder(input_dir = "../outputs/mixLL-outputs/", nsp=52, startyr=1, stopyr=77, iter=9999)
# Run the lambda_interpolation method on this data
lambdaIndicator_all_1TP <- lambda_indicator(lambdaArray_all_1TP)
#write.csv(lambdaIndicator_all_1TP$species_change, "../outputs/growthRatesTP1.csv")


# Use function to build a three-dimensional matrix. The dimensions of this array represent species, years, and iterations.
lambdaArray_all_2TP <- arrayBuilder(input_dir = "../outputs/mixLL-outputs/", nsp=52, startyr=77, stopyr=110, iter=9999)
# Run the lambda_interpolation method on this data
lambdaIndicator_all_2TP <- lambda_indicator(lambdaArray_all_2TP)
#write.csv(lambdaIndicator_all_2TP$species_change, "../outputs/growthRatesTP2.csv")
```


Argynnis adippe and Boloria euphrosyne have not converged, lets take a closer look:

```{r non-converged-models, eval=FALSE}
# 
require(sparta)


source("./function-summarise_growthRate.R")
growthRates <- summarise_growthRate("../outputs/mixLL-outputs/", 1, 77, 77,110)
View(growthRates)
write.csv(growthRates, "../outputs/growthRates.csv")
test <- read.csv("../outputs/growthRates1.csv")


############## Habitat specialists
source("./function-plot.Occdet_G.R")
input_dir <- "../outputs/CtagWatsonCrick_results_HS/"
files <- list.files(path = paste(input_dir), ignore.case = TRUE, pattern = '\\.rds$') # list of the files to loop through

# Create empty list
p_HS <- list() 

for(i in 1:length(files)){
  # read the rds file
  out <- readRDS(file.path(input_dir, files[i])) 
  p_HS[[i]] <- plot.occDet_G(out)
}


############## Wider countryside species
source("./function-plot.Occdet_G.R")
input_dir <- "../outputs/CtagWatsonCrick_results_WCS/"
files <- list.files(path = paste(input_dir), ignore.case = TRUE, pattern = '\\.rds$') # list of the files to loop through

# Create empty list
p_WCS <- list() 

for(i in 1:length(files)){
  # read the rds file
  out <- readRDS(file.path(input_dir, files[i])) 
  p_WCS[[i]] <- plot.occDet_G(out)
}

p_HS

p_WCS

plot.occDet_Gtwo(results_Boloria_euphrosyne_ctag, results_Boloria_euphrosyne_mixLL)
plot.occDet_Gtwo(results_Pyrgus_malvae_crick, results_Pyrgus_malvae_mixLL)
```

