---
title: "07 Model Selection"
author: "Galina M. Jönsson"
date: "20/04/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

**To do list:**   

* Fix SummariseRhatDetProb-function to:   
    + Handle lists   
    + handle catLL   
* Detection probability for for catLL   
* Map records    
* Tidy obs vs pred estimate plots      
* Annotate shiny plotting code   
* Generalize first Py summary code chunk   





# Background & standardisation

## Background

#### Here, I compare the three occupancy model formulations:   
- **Model A**: specifies that list length should be considered as a categorical variable. There are 3 classes: BNM records (excluding BMS records), UKBMS records, and NHM records.  
   
- **Model B**: specifies that list length should be considered as both a continuous variable and a categorical variable. There are 3 classes: BNM records (excluding BMS records) are considered continuous, whilst UKBMS records and NHM records are considered as categorical.  
   
- **Model C**: specifies that list length should be considered as both continuous and categorical variables with no year effect estimated for one class (NHM records). There are 3 classes: BNM records (excluding BMS records) are considered continuous, whilst UKBMS records and NHM records are considered as categorical.  



#### For each of 12 species: 
1. *Erebia epiphron*   
2. *Carterocephalus palaemon*   
3. *Melitaea athalia*   
4. *Polyommatus bellargus*   
5. *Thecla betulae*   
6. *Limenitis camilla*   
7. *Polygonia c-album*   
8. *Erebia aethiops*   
9. *Aglais urticae*   
10. *Pararge aegeria*   
11. *Lycaena phlaeas*   
12. *Pieris rapae*   




See the file '04-OccupancyTrends.Rmd' for detailed methods of species selection. In short, species were selected based on range (small, medium or large), weather they are Habitat Specialists or Wider Countryside Species, variance of the mean annual detection probability and record numbers. Species that show much uncertainty (high variance of the mean annual detection probability) are expected to benefit most from the Model C detection sub-model formulation. Note that there are no Small Range Wider Countryside Species nor any Large Range Habitat Specialist; hence, I chose four species from each of Small Range Habitat Specialist and Large Range Wider Countryside Species, respectively.   



**Small Range Habitat Specialist** (four species)   
The three species with the largest mean annual detection probability standard deviations (*Erebia epiphron*, *Carterocephalus palaemon* and *Melitaea athalia*) and the sixth (of 13) most variable species, *Polyommatus bellargus*, which has most records of all species in this group.   
   
   
   
**Medium Range Habitat Specialist** (two species)   
The species with the largest mean annual detection probability standard deviation (*Thecla betulae*) and fourth (of 16) most variable species, *Limenitis camilla*, because among the species in this group, it is the only species with recent range expansions and has among the highest record numbers.   
   
   
   
**Medium Range Wider Countryside Species** (two species)   
The largest mean annual detection probability standard deviation (*Polygonia c-album*) and the third (of 10) most variable species, *Erebia aethiops*, as it has among the fewest records of the species in this group.   
   
   
   
**Large Range Wider Countryside Species** (four species)   
The three species with the largest mean annual detection probability standard deviations (*Aglais urticae*, *Pararge aegeria* and *Lycaena phlaeas*) and the fourth (of 13) most variable species, *Pieris rapae*, as the genus of the third (*Aglais io*) is already represented in the sample.   







```{r load-model-outputs, message=FALSE, warning=FALSE}
#require(dplyr)
#require(devtools)
#devtools::install_github("rstudio/rmarkdown")
require(rmarkdown)


# Create a vector with species names to name model outputs
ModelNames <- c('Aglais_urticae',
               'Carterocephalus_palaemon',
               'Erebia_aethiops',
               'Erebia_epiphron',
               'Limenitis_camilla',
               'Lycaena_phlaeas',
               'Melitaea_athalia',
               'Pararge_aegeria',
               'Pieris_rapae',
               'Polygonia_c.album',
               'Polyommatus_bellargus',
               'Thecla_betulae')




########################################################## 
#######################  Model A  ########################
##########################################################

ModelNames_catLL <- paste(ModelNames, "catLL", sep="_") # Add '_catLL' to model names
catLL_list <- list() # Create an empty list
input_dir <- "../outputs/catLL-outputs/"  # Define input directory

# List files to loop through
files <- list.files(path = paste(input_dir), 
                    ignore.case = TRUE, 
                    pattern = '\\.rds$')

# Loop through, read the rds files, append to list and name elements
for(i in 1:length(files)){
  # read rds file and append to list
  catLL_list[[i]] <- readRDS(file.path(input_dir, files[i])) 
  # name element
  names(catLL_list)[[i]] <- paste(ModelNames_catLL[i])
}




########################################################## 
#######################  Model B  ########################
##########################################################

# Load and list model outputs. I specify files as folder contains all species
mixLL_list <- list(
  readRDS("../outputs/mixLL-outputs/results_Aglais_urticae_crick_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Carterocephalus_palaemon_ctag_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Erebia_aethiops_crick_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Erebia_epiphron_crick_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Limenitis_camilla_ctag_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Lycaena_phlaeas_crick_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Melitaea_athalia_watson_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Pararge_aegeria_watson_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Pieris_rapae_crick_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Polygonia_c_album_watson_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Polyommatus_bellargus_crick_mixLL.rds"),
  readRDS("../outputs/mixLL-outputs/results_Thecla_betulae_crick_mixLL.rds")
)

# Add model name '_mixLL' to species names
ModelNames_mixLL <- paste(ModelNames, "mixLL", sep="_")

# Name list elements accordingly
for(i in 1:length(mixLL_list)){
  names(mixLL_list)[[i]] <- paste(ModelNames_mixLL[i])
}




########################################################## 
#######################  Model C  ########################
##########################################################

ModelNames_mixLL2 <- paste(ModelNames, "mixLL2", sep="_") # Add '_mixLL2' to model names
mixLL2_list <- list() # Create an empty list
input_dir <- "../outputs/mixLL2-outputs/" # Define input directory

# List files to loop through
files <- list.files(path = paste(input_dir), 
                    ignore.case = TRUE, 
                    pattern = '\\.rds$')

# Loop through, read the rds files, append to list and name elements
for(i in 1:length(files)){
  # read rds file and append to list
  mixLL2_list[[i]] <- readRDS(file.path(input_dir, files[i])) 
  # name element
  names(mixLL2_list)[[i]] <- paste(ModelNames_mixLL2[i])
}




########################################################## 
#############  Check model A, B & C outputs ##############
##########################################################

# Check that the number of species observations and sites where the species have been collected and recorded are equal for model B and C. In other words whether the models are based on the same data set. 
for (i in 1:length(mixLL_list)) {
  if(isFALSE(mixLL_list[[i]]$species_sites == mixLL2_list[[i]]$species_sites &&
             catLL_list[[i]]$species_sites == mixLL_list[[i]]$species_sites &&
             catLL_list[[i]]$species_sites == mixLL2_list[[i]]$species_sites)){
    print("stop, species_observations differ")
  }
  if(isFALSE(mixLL_list[[i]]$species_sites == mixLL2_list[[i]]$species_sites &&
             catLL_list[[i]]$species_sites == mixLL_list[[i]]$species_sites &&
             catLL_list[[i]]$species_sites == mixLL2_list[[i]]$species_sites)){
    print("stop, species_sites differ")
  }
}




########################################################## 
##################  Rhat: Convergence ####################
##########################################################


### occupancy estimates

# Source function checking occupancy convergence i.e. Rhat >1.1?
source("./function-summarise_Rhat.R")

# If some species have not converged (converged-column == "N"), subset them
NotConvergedOccu <- subset((summarise_Rhat(append(catLL_list, 
                                              (append(mixLL_list,
                                              mixLL2_list))), 
                                       verbose=FALSE,
                                       is.list=TRUE)), # Specify its a list
                          grepl("N", converged))


# Source function checking detection probability convergence i.e. Rhat >1.1?
source("./function-SummariseRhatDetProb.R")

# Run function
#RhatDet_catLL <- SummariseRhatDetProb("../outputs/catLL-outputs/", 
#                              verbose=FALSE)


# If some species have not converged (converged-column == "N"), subset them
#NotConvergedDetProb_catLL <- subset(RhatDet_mixLL, grepl("N", RhatDet_catLL[,2:5]))
NotConvergedDetProb_mixLL <- subset((SummariseRhatDetProb("../outputs/mixLL-outputs/", 
                              verbose=FALSE)), 
                              grepl("N", (SummariseRhatDetProb("../outputs/mixLL-outputs/", 
                              verbose=FALSE))[,2:5]))
NotConvergedDetProb_mixLL2 <- subset((SummariseRhatDetProb("../outputs/mixLL2-outputs/", 
                              verbose=FALSE)), 
                              grepl("N", (SummariseRhatDetProb("../outputs/mixLL2-outputs/", 
                              verbose=FALSE))[,2:5]))


# If any of the model outputs have occupancy estimates that have not converged (Rhat >1.1), display them as a table.
if(nrow(NotConvergedOccu) > 0){
  knitr::kable(NotConvergedOccu)
}
if(nrow(NotConvergedDetProb_mixLL) > 0){
  knitr::kable(NotConvergedDetProb_mixLL)
}
if(nrow(NotConvergedDetProb_mixLL2) > 0){
  knitr::kable(NotConvergedDetProb_mixLL2)
}
```


## Visualisation




### Record numbers
```{r plot-record-numbers, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Load required packages
require(dplyr)
require(plyr)
require(ggplot2)
require(viridis)
require(readr)

# Load spp_vis data (any would do), which contain all visits and whether or not each of the three species have been observed at that visit
spp_vis <- read.csv('../data/formattedData/spp_vis_merged_2020-04-07.csv', header=T, na.strings=c("","NA"))

# Load occDetdata giving each visits' site ID (grid cell), list length (numper of species observed) and time period (in this case year)
occDetdata_merged_2020_04_07 <- read.csv('../data/formattedData/occDetdata_merged_2020-04-07.csv', header=T, na.strings=c("","NA"))

# Merge by visit
spp_vis <- merge(spp_vis, occDetdata_merged_2020_04_07, by="visit")

# Find unique visits
spp_vis <- unique(spp_vis)

# Create a species column
spp_vis$species <- as.factor(c("NA"))

### Make L column a source column
# Change all List lengths between 0 and 22222 to 1 (all these are BNM data)
spp_vis$L <- replace(spp_vis$L, spp_vis$L > 0 & spp_vis$L < 22222, 1)
# Make L a factor
spp_vis$L <- as.factor(spp_vis$L)
# Rename factor levels according to data type
spp_vis$L <- revalue(spp_vis$L, c("1"="BNM", "22222"="UKBMS", "33333"="NHCs"))
# Rename column to Source
colnames(spp_vis)[colnames(spp_vis)=="L"] <- "Source"


# Correct time period so it starts at 1900
spp_vis$year <- spp_vis$TP+1899



########################################################################

# Subset each species' records into dataframes and specify species identity
spp_vis_A <- subset(spp_vis, Aglais.urticae == "TRUE")
spp_vis_A$species <- c("italic('A. urticae')")
spp_vis_B <- subset(spp_vis, Pararge.aegeria == "TRUE")
spp_vis_B$species <- c("italic('P. aegeria')")
spp_vis_C <- subset(spp_vis, Polygonia.c.album == "TRUE")
spp_vis_C$species <- c("italic('P. c-album')")
spp_vis_D <- subset(spp_vis, Pieris.rapae == "TRUE")
spp_vis_D$species <- c("italic('P. rapae')")


# row bind each subsetted species' dataframe with the original spp_vis dataframe
spp_vis1 <- rbind(spp_vis_A, spp_vis_B)
spp_vis1 <- rbind(spp_vis1, spp_vis_C)
spp_vis1 <- rbind(spp_vis1, spp_vis_D)




########################################################################


# Subset each species' records into dataframes and specify species identity
spp_vis_Carterocephalus_palaemon <- subset(spp_vis, Carterocephalus.palaemon == "TRUE")
spp_vis_Carterocephalus_palaemon$species <- c("italic('C. palaemon')")
spp_vis_Erebia_aethiops <- subset(spp_vis, Erebia.aethiops == "TRUE")
spp_vis_Erebia_aethiops$species <- c("italic('E. aethiops')")
spp_vis_Erebia_epiphron <- subset(spp_vis, Erebia.epiphron == "TRUE")
spp_vis_Erebia_epiphron$species <- c("italic('E. epiphron')")
spp_vis_Melitaea_athalia <- subset(spp_vis, Melitaea.athalia == "TRUE")
spp_vis_Melitaea_athalia$species <- c("italic('M. athalia')")



# row bind each subsetted species' dataframe with the original spp_vis dataframe
spp_vis2 <- rbind(spp_vis_Carterocephalus_palaemon, spp_vis_Erebia_aethiops)
spp_vis2 <- rbind(spp_vis2, spp_vis_Melitaea_athalia)
spp_vis2 <- rbind(spp_vis2, spp_vis_Erebia_epiphron)




########################################################################


# Subset each species' records into dataframes and specify species identity
spp_vis_Limenitis_camilla <- subset(spp_vis, Limenitis.camilla == "TRUE")
spp_vis_Limenitis_camilla$species <- c("italic('L. camilla')")
spp_vis_Lycaena_phlaeas <- subset(spp_vis, Lycaena.phlaeas == "TRUE")
spp_vis_Lycaena_phlaeas$species <- c("italic('L. phlaeas')")
spp_vis_Polyommatus_bellargus <- subset(spp_vis, Polyommatus.bellargus == "TRUE")
spp_vis_Polyommatus_bellargus$species <- c("italic('P. bellargus')")
spp_vis_Thecla_betulae <- subset(spp_vis, Thecla.betulae == "TRUE")
spp_vis_Thecla_betulae$species <- c("italic('T. betulae')")


# row bind each subsetted species' dataframe with the original spp_vis dataframe
spp_vis3 <- rbind(spp_vis_Limenitis_camilla, spp_vis_Lycaena_phlaeas)
spp_vis3 <- rbind(spp_vis3, spp_vis_Polyommatus_bellargus)
spp_vis3 <- rbind(spp_vis3, spp_vis_Thecla_betulae)


################################################################

# Plot function
plot_speciesVisits <- function(data, title=NULL){
pp <- ggplot(data, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  facet_grid(~ species, labeller=label_parsed) + # facet grid by species
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Number of Visits", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
print(pp)
}
  

# Apply plot function to the three species lists
mapply(data = list(spp_vis1, 
                  spp_vis3, 
                  spp_vis2),
       FUN = plot_speciesVisits)

```







### DIC
```{r DIC, cache=TRUE, message=FALSE, echo=FALSE}
# Create empty data frame to populate
summary_out <- data.frame(Species = as.character(),
                          Best_fit = as.character(),
                          Model_B = as.numeric(),
                          Model_C = as.numeric(),
                          Difference = as.numeric())

# loop through BUGSoutput for each species pair in turn
for (i in 1:length(mixLL_list)) { # If Model B's DIC is lower than C's
  if(mixLL_list[[i]]$BUGSoutput$DIC < mixLL2_list[[i]]$BUGSoutput$DIC) {
    Best <- "B"    # Name 'Best' B
  } else {
      Best <- "C"    # Else name 'Best' C
  }
  summary_out[i, "Species"] <- paste(paste("*", spp_names[i], sep = ""), "*", sep = "")     # Populate species_name column from spp_names list and surround by asterisks
  summary_out[i, "Best_fit"] <-  Best     # Populate Best_fit column with lowest DIC model
  summary_out[i, "Model_B"] <-  mixLL_list[[i]]$BUGSoutput$DIC     # Enter Model B's DIC value
  summary_out[i, "Model_C"] <-  mixLL2_list[[i]]$BUGSoutput$DIC     # Enter Model C's DIC value
  summary_out[i, "Difference"] <- abs(mixLL_list[[i]]$BUGSoutput$DIC - mixLL2_list[[i]]$BUGSoutput$DIC)     # Enter the difference between Model B's and C's DIC values
  }

knitr::kable(summary_out)
```
**Reminder**: DIC differences are 'absolute', not relative to the DIC estimates themselves (e.g. a difference of 2 is the same thing when models' DICs are 80k and 79998 vs 2 and 4?)


# Posterior Predictive Checks

## Extract and summarise Py[j] posterior distribution

> (1)   
For each of *v* visits in each species’ model, extract 99 samples from the posterior distribution of the probability that an observation was made on that visit. In practical terms, this probability is the product of the true (unknown) occupancy, *z*~*it*~, and the detection probability, *p*~*itv*~


As the model formulation is coded in the following way:   
```{r model-example, cache=TRUE, message=FALSE, eval=FALSE}
for(j in 1:nvisit) {   
y[j] ~ dbern(Py[j])   
y.new[j] ~ dbern(Py[j])   
Py[j] <- z[Site[j],Year[j]]*p[j]   
logit(p[j]) <-  alpha.p[Year[j]] + LL.p * logL[j] + dtype2.p * DATATYPE2[j] + dtype3.p * DATATYPE3[j] }   
}  
```
   
   
the following line gives the product of the true occupancy, *z*~*it*~, and the detection probability, *p*~*itv*~
```{r py-demonstration, cache=TRUE, message=FALSE, eval=FALSE}
Py[j] <- z[Site[j],Year[j]]*p[j]   
```   

In other words, we want to extract Py[j]

> (2)   
Use each of the 99 sets of probabilities to sample a vector of potential observations under the model, by treating each visit as a potential Bernoulli trial. These vectors are 99 realizations of *y*~rep~

> (3)   
For each realization *y*~rep~ and for each year, calculate the annual proportion of sites in which the species was recorded; denote this proportion, for species *s* and year *t*, by *T*~*st*~(*y*~rep~)


```{r PostPredChecks1, eval=FALSE, message=FALSE}
require("rjags")
require("coda")
require("reshape2")
require("dplyr")
require("sparta")
require("lattice")
require("LearnBayes")
require("R2jags")


Thecla_betulae.catLL <- catLL_list[[12]]


###########################
############ 1 ############
###########################


# recompile the model
Thecla_betulae.catLL$model$recompile()

#out$model$recompile()

# Sample 100 iterations of each of the 3 chains and thin by 3, giving 99 samples
samp <- rjags:::coda.samples(model=Thecla_betulae.catLL$model, 
                             parallel=TRUE, n.cores=3,
                             variable.names="Py", 
                             n.iter=100, # Define number of iterations to run
                             thin=3) # Define thinning


###########################
############ 2 ############
###########################

### Create vectors that are are 99 realizations of yrep

# Use the 99 sets of probabilities & sample a vector of simulated observations under the model, by treating each visit as a potential Bernoulli trial. Here, we use rbinom as Bernoulli trial 
Y_rep <- lapply(samp, function(x) apply(x, 1:2, rbinom, n=1, size=1))


### Tidy the 99 realizations of yrep       

# What does this do? 
Y_rep <- melt(Y_rep)

# Tidy VisitID
Y_rep$VisitID <- as.character(gsub(Y_rep$Var2, pa="Py\\[", repl=""))
Y_rep$VisitID <- as.numeric(gsub(Y_rep$VisitID, pa="\\]", repl=""))

# Tidy iteration
Y_rep$iter <- with(Y_rep, Var1 + max(Var1)*(L1 - 1))

# Include year, siteID and the actual observation (y)        
# load the raw data
#temp_data <- out$model$data()
temp_data <- Thecla_betulae.catLL$model$data()
temp_data <- as.data.frame(with(temp_data, cbind(Site, Year, y)))
temp_data$VisitID <- 1:nrow(temp_data)
Y_rep <- merge(Y_rep[,-c(1,2,4)], temp_data)
        
#with(Y_rep, table(y,value))


###########################
############ 3 ############
###########################
        
# next calculate the reporting rate in differing ways
# first, for each site:year combination (i.e., yrep) report the number of sites for which a positive record was made
RR1 <- Y_rep %>% group_by(Site, Year, iter) %>%
  summarise(obs = max(y),sim = max(value)) %>%
  ungroup()
# the data in RR1 refer to whether the species was recorded in each site in each year
# save this file
write.csv(RR1, "../outputs/posteriorPredictiveChecks-outputs/Thecla_betulae.catLL_RR1.csv")

# next, the mean of these across years (i.e., Tst(yrep))
RR2 <- RR1 %>% group_by(Year, iter) %>%
          summarise(obs_meanSitesPerYear = mean(obs),
                    sim_meanSitesPerYear = mean(sim)
          ) %>%
          ungroup()

# the data in RR2 refer to the proportion of Sites with positive records per Year
# save this file
write.csv(RR2, "../outputs/posteriorPredictiveChecks-outputs/Thecla_betulae.catLL_RR2.csv")
```





> (4)
Calculate *T*~*gt*~(*y*~rep~) for each replicate dataset as the mean of *T*~*st*~(*y*~rep~) across species in each taxonomic group, *g*.

Here, I will calculate the mean of *T*~*st*~(*y*~rep~) across species for each model formulation, not taxonomic group (still it calling *g*).


> (5)
Calculate the mean *m* across years as *T*~*gm*~(*y*~rep~) for each replicate dataset


```{r read-PostPredCheck-Data, message=FALSE, warning=FALSE}
library(R2jags)
library(reshape2)
library(sparta)
library(LearnBayes)
library(lattice)
library(dplyr)
library(tidyr)
library(ggplot2)

# Define 99% CIs
q <- c(0.005, 0.995)

# Specify directory
path_in <- "../outputs/posteriorPredictiveChecks-outputs/"

# List relevant files in directory
spp_files <- paste(path_in, list.files(path=path_in, 
                                       pattern="\\RR2.csv$",
                                       recursive = TRUE), sep="/")
# read relevant files
data <- lapply(spp_files, read.csv)

# how many models?
print(length(data))

#append the species names
sppnames <- gsub(spp_files, patt = paste0(path_in, "/"), repl="")
sppnames <- gsub(sppnames, patt = "_RR2.csv", repl="")
sppnames <- sub(sppnames, patt = "_", repl=".")
names(data) <- sppnames


# coerce to a single dataframe
sppMeanSites <- melt(data, id=1:5)

# separate Model and Species
sppMeanSites <- sppMeanSites %>% 
  separate(L1, into=c("Species", "Model"), 
           sep="_", remove=TRUE)

# Assign correct years
sppMeanSites$Year <- sppMeanSites$Year + 1899

# Correct species names and include asteriks for italising 
sppMeanSites$Species <- gsub(sppMeanSites$Species, patt = "c.album", repl="c-album")
sppMeanSites$Species <- gsub(sppMeanSites$Species, patt = "[.]", repl=" ")
sppMeanSites$Species <- paste("*", sppMeanSites$Species, sep = "")
sppMeanSites$Species <- paste(sppMeanSites$Species, "*", sep = "")
```


Now we can calculate three other summary statistics:
* sppVarSites: the variance across years in the proportion of sites with a record
* modelMeanSites: the annual mean (across species) in the proportion of sites with a record per model
* modelVarSites: the variance across years in the above



> (6)
Calculate *T*~*gvar*~(*y*~*rep*~) as the variance across years in *T*~*gt*~(*y*~rep~) for each replicate dataset.

> (7)
Calculate the observed mean proportion of sites with records, *T*~*gm*~(*y*), and the variance across years, *T*~*gvar*~(*y*), for each replicate dataset.


```{r varianceAndModel, message=FALSE, warning=FALSE}
# now the variance across years 
sppVarSites <- sppMeanSites %>% dplyr:::group_by(Model, Species, iter) %>%
  dplyr:::summarise(obs_varSitesPerYear = var(obs_meanSitesPerYear),
            sim_varSitesPerYear = var(sim_meanSitesPerYear)
  ) %>%
  dplyr:::ungroup()

# now summarise these stats to model level
# We'll calculate the mean proportion of sites with records for species in each taxon
modelMeanSites <- sppMeanSites %>% dplyr:::group_by(Model, Year, iter) %>%
  dplyr:::summarise(obs_meanSitesPerYear = mean(obs_meanSitesPerYear),
            sim_meanSitesPerYear = mean(sim_meanSitesPerYear)
  ) %>%
  dplyr:::ungroup()

# finally, the variance among years
modelVarSites <- modelMeanSites %>% dplyr:::group_by(Model, iter) %>%
  dplyr:::summarise(obs_varSitesPerYear = var(obs_meanSitesPerYear),
            sim_varSitesPerYear = var(sim_meanSitesPerYear)
  ) %>%
  ungroup()


#####################################
sppMeanSitesYear <- sppMeanSites %>% dplyr:::group_by(Model, Species, Year) %>%
  dplyr:::summarise(obs_meanSitesPerYear = mean(obs_meanSitesPerYear),
            sim_MeanSitesPerYear = mean(sim_meanSitesPerYear),
            sim_lower = quantile(sim_meanSitesPerYear, q[1]),
            sim_upper = quantile(sim_meanSitesPerYear, q[2])
  ) %>%
  dplyr:::ungroup()

sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
```


> (8)
Summarize the distributions of *T*~*gm*~(*y*) and *T*~*gvar*~(*y*) as the mean and 95% credible intervals to demonstrate the variation in summary measures that can reasonably be expected under the model.

I will use 99% CIs as the lower 95% CI overlaps zero. 

Summarise the data, including the Bayesian *p* value

```{r BayesianP, message=FALSE, warning=FALSE}
sppMeanSites_Summ <- sppMeanSites %>% 
  dplyr:::group_by(Model, Species) %>%
    dplyr:::summarise(Bp = mean(obs_meanSitesPerYear > sim_meanSitesPerYear),
              obs = mean(obs_meanSitesPerYear),
              sim_mean = mean(sim_meanSitesPerYear),
              sim_lower = quantile(sim_meanSitesPerYear, q[1]),
              sim_upper = quantile(sim_meanSitesPerYear, q[2]),
              ) %>%
  dplyr:::ungroup()


sppVarSites_Summ <- sppVarSites %>% 
  dplyr:::group_by(Model, Species) %>%
  dplyr:::summarise(Bp = mean(obs_varSitesPerYear > sim_varSitesPerYear),
            obs = mean(obs_varSitesPerYear),
            sim_mean = mean(sim_varSitesPerYear),
            sim_lower = quantile(sim_varSitesPerYear, q[1]),
            sim_upper = quantile(sim_varSitesPerYear, q[2]),
  ) %>%
  dplyr:::ungroup()


modelMeanSites_Summ <- modelMeanSites %>% 
  dplyr:::group_by(Model) %>%
  dplyr:::summarise(Bp = mean(obs_meanSitesPerYear > sim_meanSitesPerYear),
            obs = mean(obs_meanSitesPerYear),
            sim_mean = mean(sim_meanSitesPerYear),
            sim_lower = quantile(sim_meanSitesPerYear, q[1]),
            sim_upper = quantile(sim_meanSitesPerYear, q[2]),
  ) %>%
  dplyr:::ungroup()


modelVarSites_Summ <- modelVarSites %>% 
  dplyr:::group_by(Model) %>%
  dplyr:::summarise(Bp = mean(obs_varSitesPerYear > sim_varSitesPerYear),
            obs = mean(obs_varSitesPerYear),
            sim_mean = mean(sim_varSitesPerYear),
            sim_lower = quantile(sim_varSitesPerYear, q[1]),
            sim_upper = quantile(sim_varSitesPerYear, q[2]),
  ) %>%
  dplyr:::ungroup()

```


### Visualise Py summary {.tabset .tabset-dropdown}

#### Observed vs Predicted data
```{r plot-obsVsPred, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
modelMeanBp_Summ <- modelMeanSites %>% 
  dplyr:::group_by(Model) %>%
  dplyr:::summarise(Bp = mean(obs_meanSitesPerYear > sim_meanSitesPerYear),
            obs = mean(obs_meanSitesPerYear),
            sim_mean = mean(sim_meanSitesPerYear),
            sim_lower = quantile(sim_meanSitesPerYear, q[1]),
            sim_upper = quantile(sim_meanSitesPerYear, q[2]),
  ) %>%
  dplyr:::ungroup()

plot_PredPostTest <- function(data, title=NULL){
  gp <- ggplot(data = data,) +
    geom_point(aes(x=obs, y=sim_mean, col=Model), alpha=0.7) +
    geom_errorbar(aes(x=obs, ymin=sim_lower, ymax=sim_upper, col=Model), 
               alpha=0.7) +
    geom_abline(aes(intercept=0, slope=1)) + 
    scale_x_log10() + scale_y_log10() + 
    xlab("Observed from data") + ylab("Model prediction") + 
    ggtitle(title) + theme_bw()
  print(gp)
}

  

mapply(data = list(sppMeanSites_Summ, 
                  sppVarSites_Summ), # 
                  #modelMeanSites_Summ,
                  #modelVarSites_Summ), 
        title = list("Mean (across years) proportion of sites with a record - 99%CIs", 
                  "Variance (across years) in proportion of sites per year - 99%CIs"), #, 
                  #"Mean proportion of sites per year (mean across species) - 99%CIs",
                  #"Variance across years in proportion of sites per year across species - 99%CIs"),   
       FUN = plot_PredPostTest)
```




#### Bayesian *p*: Model A vs Model B


Now plot the Bayesian *p*-values (the proportion of simulated values that are lower than the observed; the 'best' Bayesian *p*-value is 0.5) 
```{r plot-Bp-AvsB, echo=FALSE, warning=FALSE, cache=TRUE}
require(tidyr)
require(viridis)
require(ggtext)

ModelMeanBp_Summ <- spread((sppMeanSites_Summ[,1:3]), Model, Bp)
ModelVarBp_Summ <- spread((sppVarSites_Summ[,1:3]), Model, Bp)

# Rename
names(ModelVarBp_Summ)[2] <- "catLL_var"
names(ModelVarBp_Summ)[3] <- "mixLL_var"
names(ModelVarBp_Summ)[4] <- "mixLL2_var"

ModelBp_Summ <- data.frame(ModelMeanBp_Summ)
ModelVarBp_Summ <- data.frame(ModelVarBp_Summ)
ModelBp_Summ[,5:7] <- data.frame(ModelVarBp_Summ[,2:4])


ggplot(ModelBp_Summ, aes(x=catLL, y=mixLL, group=Species)) + 
  geom_point(aes(shape=Species, color=Species),  size=5) +
  scale_color_manual(values =c (viridis(15))) +
  scale_shape_manual(values = c(15:18, 21:24, 15:18, 21:23)) + 
  ggtitle("Bayesian p-values",
          subtitle = "Mean proportion of sites with records across years") + 
  theme_bw() +
  geom_abline(aes(intercept=0, slope=1)) + 
  xlab("Model A") + ylab("Model B") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  geom_vline(xintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  theme(legend.text = element_markdown()) +
  theme(text=element_text(size=13)) 


ggplot(ModelBp_Summ, aes(x=catLL_var, y=mixLL2_var, group=Species)) +
  geom_point(aes(shape=Species, color=Species),  size=5) +
  scale_color_manual(values =c(viridis(15))) +
  scale_shape_manual(values = c(15:18, 21:24, 15:18, 21:23)) + 
  ggtitle("Bayesian p",
          subtitle = "Variance proportion of sites with records across years") + 
  theme_bw() +
  geom_abline(aes(intercept=0, slope=1)) + 
  xlab("Model A") + ylab("Model B") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  geom_vline(xintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  theme(legend.text = element_markdown()) +
  theme(text=element_text(size=13)) 
```
  


#### Bayesian *p*: Model A vs. Model C
```{r plot-Bp-AvsC, echo=FALSE, warning=FALSE, cache=TRUE}
require(tidyr)
require(viridis)
require(ggtext)

ggplot(ModelBp_Summ, aes(x=catLL, y=mixLL2, group=Species)) + 
  geom_point(aes(shape=Species, color=Species),  size=5) +
  scale_color_manual(values =c(viridis(15))) +
  scale_shape_manual(values = c(15:18, 21:24, 15:18, 21:23)) + 
  ggtitle("Bayesian p-values",
          subtitle = "Mean proportion of sites with records across years") + 
  theme_bw() +
  geom_abline(aes(intercept=0, slope=1)) + 
  xlab("Model A") + ylab("Model C") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  geom_vline(xintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  theme(legend.text = element_markdown()) +
  theme(text=element_text(size=13)) 

ggplot(ModelBp_Summ, aes(x=catLL_var, y=mixLL2_var, group=Species)) +
  geom_point(aes(shape=Species, color=Species),  size=5) +
  scale_color_manual(values =c(viridis(15))) +
  scale_shape_manual(values = c(15:18, 21:24, 15:18, 21:23)) + 
  ggtitle("Bayesian p",
          subtitle = "Variance proportion of sites with records across years") + 
  theme_bw() +
  geom_abline(aes(intercept=0, slope=1)) + 
  xlab("Model A") + ylab("Model C") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  geom_vline(xintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  theme(legend.text = element_markdown()) +
  theme(text=element_text(size=13)) 
```


#### Bayesian *p*: Model B vs. Model C
```{r plot-Bp-BvsC, echo=FALSE, warning=FALSE, cache=TRUE}
require(tidyr)
require(viridis)
require(ggtext)

ggplot(ModelBp_Summ, aes(x=mixLL, y=mixLL2, group=Species)) + 
  geom_point(aes(shape=Species, color=Species),  size=5) +
  scale_color_manual(values =c(viridis(15))) +
  scale_shape_manual(values = c(15:18, 21:24, 15:18, 21:23)) + 
  ggtitle("Bayesian p-values",
          subtitle = "Mean proportion of sites with records across years") + 
  theme_bw() +
  geom_abline(aes(intercept=0, slope=1)) + 
  xlab("Model B") + ylab("Model C") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  geom_vline(xintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  theme(legend.text = element_markdown()) +
  theme(text=element_text(size=13)) 


ggplot(ModelBp_Summ, aes(x=mixLL_var, y=mixLL2_var, group=Species)) +
  geom_point(aes(shape=Species, color=Species),  size=5) +
  scale_color_manual(values =c(viridis(15))) +
  scale_shape_manual(values = c(15:18, 21:24, 15:18, 21:23)) + 
  ggtitle("Bayesian p",
          subtitle = "Variance proportion of sites with records across years") + 
  theme_bw() +
  geom_abline(aes(intercept=0, slope=1)) + 
  xlab("Model B") + ylab("Model C") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  geom_vline(xintercept=0.5, linetype="dashed", color = "red", alpha=0.5) +
  theme(legend.text = element_markdown()) +
  theme(text=element_text(size=13)) 
```
**Notes**   

* Model B/mixLL: Six of 12 species have better p-values   
* Model C/mixLL2: Six of 12 species have better p-values   
* DIC-values favor Model B for all species but three: *M athalia*, *C. palaemon* (Model C better according to p-values) and *L. phlaeas* (p-values practically identical)    
* Two species have practically equal p-value:  
    + *L. phlaeas*: some differences in occupancy trends but this species shows cyclical    occupancies over time and both models capture the cyclical pattern     
    + *P. aegeria*: Practically identical occupancy trends     
* *Erebia epiphron*  shows the largest difference in p-vale between the two models    
    + Occupancy estimates for the two models are practically identical       
    + This species has fewest BMS (i.e. standardized) records       
    + This species has fewest NHC records       
* *Limenitis camilla*  shows the second largest difference in p-vale between the two models (0.48 for mixLL2/Model C and 0.57 for for mixLL/Model B)     
    + Occupancy estimates for the two models are practically identical (although much underestimated in model A)       


**Means Bayesian-p: Model B vs C**

1. *Erebia epiphron*: C much better     
2. *Carterocephalus palaemon*: C better    
3. *Melitaea athalia*: C better    
4. *Polyommatus bellargus*: B better    
5. *Thecla betulae*: B slightly better    
6. *Limenitis camilla*: C much better    
7. *Polygonia c-album*: B better    
8. *Erebia aethiops*: C better    
9. *Aglais urticae*: B better    
10. *Pararge aegeria*: practically identical    
11. *Lycaena phlaeas*: practically identical     
12. *Pieris rapae*: B slightly better      


**Nick**   
   
* Given that the 'optimal' p-value is 0.5, is *Limenitis camilla* (0.48 for mixLL2/Model C and 0.57 for for mixLL/Model B) "more" interesting than let's say *Erebia epiphron* (which shows the largest difference in p-vale between the two models)?    Two questions: which model is better ( for most id doesnt make a difference but for E. epophron)
* Variance: **very** poor p-values...
* Is the sum of of distance from 0.5 across species for model B (1.597614) vs model C (1.555207) of interest?  
* Data visualisation: colour blind-friendly vs. visually appealing... 



```{r BpDiff, echo=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
mixLL_out <- 0
mixLL2_out <- 0
for(i in 1:nrow(ModelBp_Summ)){
  mixLLtemp <- abs(0.5 - ModelBp_Summ[i,"mixLL"])
  mixLL2temp <- abs(0.5 - ModelBp_Summ[i,"mixLL2"])
  mixLL_out <- mixLL_out + mixLLtemp
  mixLL2_out <- mixLL2_out + mixLL2temp
}

for(i in 1:nrow(ModelBp_Summ)){
  mixLL2temp <- abs(0.5 - ModelBp_Summ[i,"mixLL2"])
}
```
  
  
  
```{r plotBp2, echo=FALSE, warning=FALSE, eval=FALSE}
hist(sppMeanSites_Summ$Bp)
hist(sppVarSites_Summ$Bp)
hist(modelMeanSites_Summ$Bp) ### MixLL_2 on the left
hist(modelVarSites_Summ$Bp)
```


```{r savePySummary, echo=FALSE}
# Finally save the outputs
#write.csv(taxonMeanSites_Summ, file="taxonMeanSites_Summ.csv")
#write.csv(taxonVarSites_Summ, file="taxonVarSites_Summ.csv")
```


# Root-Mean-Square Error (RMSE)
$$RMSE = \sqrt{\sum_{i=1}^{N} \frac{(Predicted_{i}-Actual_{i})^2}{N}}$$


```{r RMSE, echo=FALSE}
# Importing the required package
library(Metrics)
  
# Taking two vectors
#actual = c(1.5, 1.0, 2.0, 7.4, 5.8, 6.6)         
#predicted = c(1.0, 1.1, 2.5, 7.3, 6.0, 6.2)      
  
# Calculating RMSE using rmse()         
#result = rmse(actual, predicted)
  
# Printing the value
#print(result)
```



# Dunn-Smyth residuals

**Nick** - in contract to Wright et al, Warton uses 'absolute estimates' per visit rather than draws from the posterior distribution. Should I take the mean of (how many?) draws from the posterior distribution? 

https://www.tandfonline.com/doi/abs/10.1080/10618600.1996.10474708

Here, we plot Dunn-Smyth residuals ([Dunn & Smyth, 1997](https://www.tandfonline.com/doi/abs/10.1080/10618600.1996.10474708)) according to the methodology of [Warton *et al.* (2017)](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.12761). We only plot Dunn-Smyth (DS) residuals for one model (*A. urticae* model B) to demonstrate the plotting procedure, especially given that our data is formatted differently to that of Warton *et al.*, and we wish to plot residuals for multi-year occupancy models, whilst the authors used a single season example; hence, we had to make some alterations to the functions provided by the authors (namely residuals.occMod() and DS.resid.plot(); Supplementary Data 6 of Warton *et al.*, 2017). Dunn-Smyth residuals for all species and models can be found under 'species-wise summary'.    
 
   
   
   
Examples of papers that have produced DS residuals for multi-year occupancy models:   

* https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0245973#pone-0245973-g004   
* https://doi.org/10.1016/j.biocon.2019.03.010   
* https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6509396/   


## Equations for residuals 
    

**DS detection residual cumulative distribution function**, $F_{det}$, use the total number of detections at a site ($i$), denoted $Y_{i}$, conditional on there having been detections ($F_{det}(Y_{i}|Y_{i}>1)$). 


> If the detection probability is not constant across surveys[, $F_{det}(Y_{i})$] is computed as a summation across all possible [detection] histories* with $Y_{i}$ or fewer detections. The cumulative distribution function of $Y_{i}$ can be written as:   

*Recall: all possible detection histories to site $i$, which has been vistited ($v$) three times within year $t$ would be $2^3 = 8$ (2 for the two options present (1) or absent (0), and 3 for the number of visits, $v$): 010, 000, 001, 011, 111, 110, 100, 101   

$$F_{det}(Y_{i}|Y_{i}>1) =\left( \frac{1}{1-(1 - \prod_{v=1}^{V}(1-p_{iv})} \right)\sum_{x|1≤Y_{x}≤Y_{i}}^{}  \prod_{v=1}^{V}-p_{iv}^{xv}(1-p_{iv})^{1- _{xv}}$$
where $Y_x=\sum_{v=1}^{V} x_v$ is the total number of detections (i.e, 1s) in detection history $x$. In other words, the equation loops through all possible detection histories, $x$, (i.e. $2^V$) to site $i$, given that there has been detections ($F_{det}(Y_{i}|Y_{i}>1)$). This can become computationaly expensive when the total number or visits, $V$, and the total number of detections at a site, $Y_{i}$, aren't small since the possible number of detection histories, $x$, with $Y_{i}$ or fewer detection becomes big (e.g., 4 visits to site $i$ of which 2 were detections gives cumuulative probability 0.6875 & 16($2^4$) * 0.6875 == **11** possible detection histories, $x$, whilst 15 visits to site $i$ of which 7 were detections gives cumulative probability 0.69638061523 & **22819** possible detection histories).     
   
   
   
**In other words, what I need is:**    
* The total number of detections at a site ($i$), $Y_{i}$ so the parameter y[j]. Then I'll need to sum the number of visits (j) to the same site within the same year      
* The detection probability $p_{iv}$ at visit $v$ to site ($i$) so the parameter p[j]. Then I'll need to sum the number of visits (j) to the same site within the same year      



    
**The occupancy residual cumulative distribution function**, $F_{occ}$, uses a key piece of information: whether or not at least one detection has occured at site $i$ ($j_{i} = 1$)  or not ($j_{i} = 0$). If $j_{i} = 0$, the cumulative distribution function for a site is not soley a function of occupancy probability ($\psi_{i}$), but also detection probability ($p_{iv}$), and is expressed as following for site $i$:   


$$
F_{occ}(j_{i}) = \begin{cases} 1-\psi_{i} + \psi_{i} \prod_{v=1}^{V}- (1-p_{iv}) \ \ \ \ \ \ \ \ \ \ {\sf if} \ j_{i} = 0 \\1\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\sf if} \ j_{i} = 1\end{cases}

$$
      
      
      
      
Hence, we need to extract the following parameters from our models to calculate occupancy and detection residual cumulative distribution functions:    
* Probability of occurrence $\psi_{i}^{[n]}$ = 'muZ'   
* Detection probability $p_{iv}^{[n]}$ = 'p'   
* True occupancy $z_{i}^{[n]}$ = 'z'   
* Estimated occupancy $y_{i}^{[n]}$ = 'y'         
      
      
      
      
**Finally**, both DS detection residual cumulative distributions and occupancy residual cumulative distributions are transformed using the below equation, which introduces a bit of random noise to generate residuals from discrete binomial variables (occupancy) with a bit of "jitter" to remove their discreteness. Additionally, Dunn-Smyth proposed mapping these onto the standard normal distribution so that hey can be interpreted like residuals from a regular linear regression.   

    
$$
\Phi(z) = (1- u)F(x) + uF\_(x)
$$
where $\Phi(\cdot)$ is the cumulative distribution function of a standard normal variable, and $F(\cdot)$ is the cumulative distribution function found from the two above equations ($F_{det}(Y_{i}|Y_{i}>1)$ and $F_{occ}(j_{i})$). $F\_(\cdot)$ is the previous value of $F(\cdot)$ **????? FROM THE PREVIOUS SITE??** and $u$ is a randomly generated value from the standard uniform distribution (**????? EQUALLY LIKELY TO TAKE ANY VALUE BETWEEN 0 AND 1????**)








# Wright residuals
    
    
    
Here, we plot residuals according to the methodology of [Wright *et al.*, (2019)](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1002/ecy.2703). We only plot Wright (W) residuals for one model output (*A. urticae* model B) to demonstrate the procedure, especially given that we wish to plot residuals for multi-year occupancy models, whilst the authors used single season examples. Wright residuals for all species and models can be found under 'species-wise summary'.    
   

    

## Equations for residuals
   
The occupancy residuals for site $i$ and iteration $n$ is expressed as: 
$$o_{i}^{[n]} = z_{i}^{[n]} - \psi_{i}^{[n]}$$
Where $z_{i}^{[n]}$ is drawn from the posterior distribution of the occupancy state at site $i$.   
    
    
The detection residuals also use the posterior distribution of the occupancy state. For posterior iteration $n$, the detection residuals for site $i$ and visit $v$ is expressed as: 
$$[y_{iv}^{[n]}|z_{i}^{[n]} = 1] = y_{iv} - p_{iv}^{[n]}$$

> These detection residuals are only defined for “occupied” sites based on the latent state $z_{i}^{[n]}$ for a given posterior draw $n$. By conditioning on a posterior draw of the occupancy state ($z_{i}^{[n]}$), sites without any detection ($y_{i}$ containing all zeros) still contribute detection residuals for some posterior draws because these sites might actually be occupied ($z_{i}^{[n]} = 1$ for some iterations. Additionally, the total number of detection residuals will vary across posterior draws because of the uncertainty in true occupancy at sites with no detections. 

from [Wright *et al.*, (2019)](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1002/ecy.2703)
   
   
   
**Note** that the above equations are for single-season occupancy models.
   
    
#### Defining residuals for occupancy models requires:   
* Estimates of probability of occurrence $\psi_{i}^{[n]}$ per site ($i$). Furthermore, as we used a multi-season occupancy model, we further need to separate estimates of probability of occurrence per survey **and** year.   
* Estimates of detection probability $p_{iv}^{[n]}$ per survey ($v$) and site ($i$). Again, as we used a multi-season occupancy model, we further need to separate detection probability estimates per survey **and** year.  
* Estimates of true occupancy $z_{i}^{[n]}$ per site ($i$). Furthermore, as we used a multi-season occupancy model, we further need to separate estimates of probability of occurrence per survey **and** year.   
    
    
## Our model formulation
   
### State submodel
    
The true occupancy, $z_{it}$, at site $i$ in year $t$ has a Bernoulli distribution governed by the probability of occurrence $\psi_{it}$: 
$$z_{it} \sim {\sf Bernoulli}(\psi_{it})$$
where the logit of the probability of occurrence $\psi_{it}$ varies between sites and years:
$${\sf logit}(\psi_{it}) = b_{t} + u_{i}$$
where $b_{t}$ and $u_{i}$ denote the year and site effects, respectively
    
    
    
The JAGS code for the state submodel is as follows:  
```{r seeStateJAGScode, eval=FALSE, echo=TRUE}
for (i in 1:nsite){ 
  for (t in 1:nyear){   
    z[i,t] ~ dbern(muZ[i,t]) 
    logit(muZ[i,t])<- a[t] + eta[i] 
  }}
```
Hence, we want to extract the parameter 'muZ', denoting the probability of occurrence, $\psi_{it}$, and 'z', denoting the true occupancy $\psi_{it}. 
   
   
   
### Detection submodel
   
The true occupancy state $z_{iv}$, at site $i$ at visit $v$ is conditional on whether there has been an observation of the species ($y_{iv}$; i.e. the binary data) and is described as being drawn from a Bernoulli distribution governed by the detection probability, $p_{itv}$, and is conditional on the occupancy state ($z_{iv}$; i.e. the species being present at visits $v$ to site $i$)
$$y_{iv}|z_{iv} = {\sf Bernoulli}(p_{itv} * z_{iv})$$
   
   
Here, we use three formulations of the detection submodel. However, what we are interested in for all three is the detection probability, $p_{itv}$, at site $i$ in year $t$ at visit $v$. Below 
    
    
    
    
##### Model A
   
where the logit of the detection probability, $p_{itv}$ varies between sites ($i$), years ($t$) and visits ($v$):
$${\sf logit}(p_{itv}) = a_{t} + \beta_{1} * {\sf BMSdata}_{itv} +  \beta_{2} * {\sf NHCdata}_{itv} $$
where $a_{t}$ is the year effect (and detection probability of BNM data), and $\beta_{1}$ and $\beta_{2}$ are the effects of BMS and NHC visits ($v$) to site ($i$) in year ($t$).    
    
    
    
The JAGS code for the observation submodel is as follows: 
```{r seeObservationJAGScode, eval=FALSE, echo=TRUE}
for(j in 1:nvisit) {
  y[j] ~ dbern(Py[j])
  Py[j]<- z[Site[j],Year[j]]*p[j]
  logit(p[j]) <-  alpha.p[Year[j]] + LL.p*logL[j] + dtype2.p*DATATYPE2[j] + dtype3.p*DATATYPE3[j]
} }
```
Hence, we want to extract the parameters 'p' and 'Py', denoting the detection probability, $p_{itv}$ and the true occupancy state $z_{iv}$ for visit $v$ to site $i$ (here denoted 'j').   
    
    
    
## Example *Aglais urticae*-Model B
    
We need to extract the following parameters from our models:    
* Probability of occurrence $\psi_{i}^{[n]}$ = 'muZ'   
* Detection probability $p_{iv}^{[n]}$ = 'p'   
* Estimated true occupancy $z_{i}^{[n]}$ = 'z'   
* Detection histories (1 or 0; detected or not) $y_{i}^{[n]}$ = 'y'   
    
    
    
Here, we show how we plot Wright residuals using *Aglais urticae*-Model B outputs as an example. We will use 99 iterations and start by ignoring year (i.e. plot residuals from all years on one plot)
```{r exampleWrightResiduals, echo=TRUE, eval=FALSE}
# Load required packages
require("rjags")
require("coda")
require("reshape2")
require("dplyr")
require("sparta")
require("lattice")
require("LearnBayes")
require("R2jags")
require("data.table")

# Read the model output
Aglais_urticae_mixLL <- readRDS("../outputs/mixLL-outputs/results_Aglais_urticae_crick_mixLL.rds")

# recompile the model
Aglais_urticae_mixLL$model$recompile()

# Sample 100 iterations of each of the 3 chains and thin by 3, giving 99 samples of "muZ", "z", and "p"
samp <- rjags:::coda.samples(model=Aglais_urticae_mixLL$model, 
                             parallel=TRUE, n.cores=3,
                             variable.names=c("muZ", "z", "p"), 
                             n.iter=100, # Define number of iterations to run
                             thin=3)
# This produced a mcmc.list of length three: one list per chain with column names in the format muZ[i,t], z[i,t] and p[j] (where j denotes visit; v) with a total length of 1366187 columns for p[j], and 319000 columns for muZ[i,t] and z[i,t], of samples estimates and 33 rows, each, totalling 99 samples 




### Separate columns station with "muZ", "z" and "p" (will also turn the mcmc.list to a 'regular' list)
# Define chain number
chainID <- c(1,2,3)

### As 'regular' lists, we can use the data.table-function melt()
# recall the format p[j] vs muZ[i,t] and z[i,t] (where j denotes visit)
# This makes p straight forward: 
p <- lapply(samp, function(x) x[, grepl("p", dimnames(x)[[2]])])

# Add Chain number to list
#1
#p2 <- mapply(cbind, p, "chain"=chainID, SIMPLIFY=F)
#2
#p2<-lapply(p, function(x) 
#  cbind(x, chain = chainID))
#3
#p[[1]]$chain <- rep(1, 99)
#p[[2]]$chain <- rep(2, 99)
#p[[3]]$chain <- rep(3, 99)

p <- rbindlist(lapply(p, as.data.frame))
p <- melt(p)

#### However, we need to separate the years in muZ[i,t] and z[i,t] (where i denotes site, and t year)
muZ <- lapply(samp, function(x) x[, grepl("muZ", dimnames(x)[[2]])])
muZ <- rbindlist(lapply(muZ, as.data.frame))
muZ <- melt(muZ)

# Separate the first year
#muZ_yr1 <- lapply(muZ, function(x) x[, grepl(",1]", dimnames(x)[[2]])])
#muZ_yr1 <- rbindlist(lapply(muZ_yr1, as.data.frame))
#muZ_yr1 <- melt(muZ_yr1)

z <- lapply(samp, function(x) x[, grepl("z", dimnames(x)[[2]])])
z <- rbindlist(lapply(z, as.data.frame))
z <- melt(z)
# Separate the first year
#z_yr1 <- lapply(z, function(x) x[, grepl(",1]", dimnames(x)[[2]])])
#z_yr1 <- rbindlist(lapply(z_yr1, as.data.frame))
#z_yr1 <- melt(z_yr1)



#### Tidy data

# Tidy VisitID column for p
p$VisitID <- as.character(gsub(p$variable, pa="p\\[", repl=""))
p$VisitID <- as.numeric(gsub(p$VisitID, pa="\\]", repl=""))

# Tidy year and site columns for muZ and z
muZ$Site <- as.character(gsub(muZ$variable, pa="muZ\\[", repl="")) 
muZ$Site <- as.numeric(gsub(muZ$Site, pa="\\,[0-9]*\\]", repl="")) # 2750 
muZ$Year <- as.character(gsub(muZ$variable, pa="muZ\\[[0-9]*\\,", repl=""))
muZ$Year <- as.numeric(gsub(muZ$Year, pa="\\]", repl="")) # 116 
z$Site <- as.character(gsub(z$variable, pa="z\\[", repl="")) 
z$Site <- as.numeric(gsub(z$Site, pa="\\,[0-9]*\\]", repl="")) # 2750 
z$Year <- as.character(gsub(z$variable, pa="z\\[[0-9]*\\,", repl=""))
z$Year <- as.numeric(gsub(z$Year, pa="\\]", repl="")) # 116 


# Tidy iteration
p <- p %>% group_by(variable) %>% mutate(iter = seq_along(variable))
muZ <- muZ %>% group_by(variable) %>% mutate(iter = seq_along(variable))
z <- z %>% group_by(variable) %>% mutate(iter = seq_along(variable))

# The above code output is a grouped data frame, so we convert it to a 'regular' data frame again
muZ <- as.data.frame(muZ[,c("variable", "value", "Site", "Year", "iter")])
z <- as.data.frame(z[,c("variable", "value", "Site", "Year", "iter")])
p <- as.data.frame(p[,c("variable", "value", "VisitID", "iter")])


# Remove site:year combinations where there have been no visits 
# load the raw data
temp_data <- results_Aglais_urticae_mixLL$model$data()

# subset year, siteID and the actual observation (y)
temp_data <- as.data.frame(with(temp_data, cbind(Site, Year, y)))  
temp_data$VisitID <- 1:nrow(temp_data)

# Finally, merge with p dataframe
p <- merge(p, temp_data)

```
muZ[i,t] and z[i,t] will be loner than p[j] (all combinations of sites (?????) and year (116) vs. the total numer of visits (for the species or only focal species??????))   
   
   
   
   
 
**To DO**  Double check iteratiosn are correct: I've tried three ways of adding a vector for chain number (i.e. 1,2 or 3) to the lists before melt()-ing but neither of my attempts have worked, do you reckon my shortcut for adding iteration numbers gives accurate results?     
   
   
   
   
## Wright occupancy residuals  
    
    
    
The occupancy residuals for site $i$ and iteration $n$ is expressed as: 
$$o_{i}^{[n]} = z_{i}^{[n]} - \psi_{i}^{[n]}$$
Where $z_{i}^{[n]}$ is drawn from the posterior distribution of the occupancy state at site $i$. 
   
   
Making calculating the residuals as easy as:   
```{r exampleWrightOccupancyResCalc, echo=TRUE, eval=FALSE}

# Rename value and variable 
names(muZ)[names(muZ) == "value"] <- "muZ_value"
names(z)[names(z) == "value"] <- "z_value"
names(muZ)[names(muZ) == "variable"] <- "muZ_variable"
names(z)[names(z) == "variable"] <- "z_variable"

# merge 
occuWright <- merge(muZ, z)

# Find residuals
occuWright$res <- (occuWright$z_value - occuWright$muZ_value)

# Save file
write.csv(occuWright, "../outputs/WrightResid/Aglais_urticae_occuWright.csv")
``` 
   
   
   
## Wright Detection Residuals
    
    
The detection residuals also use the posterior distribution of the occupancy state. For posterior iteration $n$, the detection residuals for site $i$ and visit $v$ is expressed as: 
$$[y_{iv}^{[n]}|z_{i}^{[n]} = 1] = y_{iv} - p_{iv}^{[n]}$$
   

   
Calculating the residuals: 
```{r exampleWrightDetectionResCalc, echo=TRUE, eval=FALSE}
# Rename value and variable 
names(p)[names(p) == "value"] <- "p_value"
names(p)[names(p) == "variable"] <- "p_variable"

# merge but specify to only include z values for sites and years where there actually was a visit ie p[j]
detWright <- merge(p, z, all.x=TRUE, all.y=FALSE)

# Check the number of estimated occupancy (i.e. 1s) 
#summary(as.factor(detWright$z_value)) 
# 131842110 1s
# 3410403 0s

# Subset the visits where estimated occupancy is 1 (as these are the visits used to find residuals)
detWright <- subset(detWright, z_value == 1) 

# Find residuals
detWright$res <- (detWright$y - detWright$p_value)

# Save file
write.csv(detWright, "../outputs/WrightResid/Aglais_urticae_detWright.csv")
``` 
     
    

## Plot binned residuals
We plot the Wright residuals as binned residuals according to the methodology of [Gelman *et al*., (2000)](https://doi.org/10.1111/1467-9876.00190), who state that binned residuals work well for *plots of binned or smoothed residuals versus predictors and specific discrepancy variables created on the basis of the particular concerns arising in an application* whilst [Wright *et al.*, (2019)](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1002/ecy.2703) state that *the values of a covariate are used to form approximately equally sized groups. Plotting the averages of the residuals in each group vs. the corresponding average covariate values can reveal unexplained structure in fitted probabilities. Examining such residual plots from multiple posterior draws characterizes variation over the posterior distribution*. Here, we use the default number of categories (bins) based on their fitted values in which the data are divided. Default=NULL and will take the value of nclass according to the $n$ such that if $n >=100$, nclass=floor(sqrt(length(x))); if $10<n<100$, nclass=10; if $n<10$, nclass=floor(n/2). However, the authors note *There is typically some arbitrariness in choosing the number of bins: each bin should contain enough points so that the averaged residuals are not too noisy, but it helps to have also many bins so as to see more local patterns in the residuals (see Gelman and Hill, Data Analysis Using Regression and Multilevel/Hierarchical Models, page 97).*       


[Gelman *et al*., (2000)](https://doi.org/10.1111/1467-9876.00190) state that quantile–quantile plots of these binned residuals does not work well, whilst [Wright *et al.*, (2019)](https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1002/ecy.2703) don't mention quantile–quantile plots of their residuals.    
   
   
**Nick**   
* What are your thoughts on quantile–quantile plots of "unbinned" residuals?    


**To do**   
* Gelman and Hill, Data Analysis Using Regression and Multilevel/Hierarchical Models, page 97    


    
    
    
    
  
```{r example-occWrightPlot, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
# Read files
occWright <- read.csv("../outputs/WrightResid/Aglais_urticae_occWright.csv")

# Load required packages
require(arm)

# Set up 2 rows 2 columns
#par(mfrow = c(2, 3))



### Plot

# It seems like the function can't handle the number of residuals since I get the following error when attempting to apply it to the whole data:
# NAs produced by integer overflowError in if (any(breaks.index == 0)) nclass <- 1 : missing value where TRUE/FALSE needed

# I therefore subset years
# 31581000/116 = 272250 observations per year

occuWright_5iter <- subset(occuWright, iter <= 5) 
occuWright_yr1to6 <- subset(occuWright, Year <= 6) 
occuWright_yr76to82 <- subset((subset(occuWright, Year >= 76)), Year < 82)


binnedplot(occuWright_yr1to6$muZ_value, 
           occuWright_yr1to6$res,
           xlab="Occupancy", 
           ylab="Binned residuals",
           main = "Binned residual plot: 1900-1905")


binnedplot(occuWright_yr76to82$muZ_value, 
           occuWright_yr76to82$res,
           xlab="Occupancy", 
           ylab="Binned residuals",
           main = "Binned residual plot: 1975-1981")

binnedplot((occuWright_5iter$Year+1899), 
           occuWright_5iter$res,
           xlab="Year", 
           ylab="Binned residuals",
           main = "Binned residual plot: 5 iterations")
# gray lines indicate plus and minus 2 standard-error bounds, within which one would expect about 95% of the binned residuals to fall, if the model were actually true.


### Q-Q plots

# occupancy residuals
#qqnorm(occuWright$res, 
#       main="Occupancy", 
#       xlab="Theoretical Quantiles", 
#       ylab="Standardised Residuals")
#qqline(occuWright$res) #add a straight diagonal line to the plot

# detection residuals
#qqnorm(detWright$res, 
#       main="Occupancy", 
#       xlab="Theoretical Quantiles", 
#       ylab="Standardised Residuals")
#qqline(detResidW$res)
```
   
   
Gray lines indicate plus and minus 2 standard-error bounds, within which one would expect about 95% of the binned residuals to fall, if the model were actually true.



```{r example-detWrightPlot, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
# Read files
detWright <- read.csv("../outputs/WrightResid/Aglais_urticae_detWright.csv")

# Load required packages
require(arm)


### Plot

# It seems like the function can't handle the number of residuals since I get the following error when attempting to apply it to the whole data:
# NAs produced by integer overflowError in if (any(breaks.index == 0)) nclass <- 1 : missing value where TRUE/FALSE needed



detWright_yr1to6 <- subset(detWright, Year <= 6) 
detWright_yr76to82 <- subset((subset(detWright, Year >= 76)), Year < 82)


binnedplot(detWright_yr1to6$p_value, 
           detWright_yr1to6$res,
           xlab="Detection Probability", 
           ylab="Binned residuals",
           main = "Binned residual plot: 1900-1905")

binnedplot((detWright_yr1to6$Year+1899), 
           detWright_yr1to6$res,
           xlab="Year", 
           ylab="Binned residuals",
           main = "Binned residual plot: 1900-1905")

binnedplot(detWright_yr76to82$p_value, 
           detWright_yr76to82$res,
           xlab="Detection Probability", 
           ylab="Binned residuals",
           main = "Binned residual plot: 1975-1981")

binnedplot((detWright_yr76to82$Year+1899), 
           detWright_yr76to82$res,
           xlab="Year", 
           ylab="Binned residuals",
           main = "Binned residual plot: 1975-1981")
# gray lines indicate plus and minus 2 standard-error bounds, within which one would expect about 95% of the binned residuals to fall, if the model were actually true.



### Q-Q plots

# occupancy residuals
#qqnorm(occuWright$res, 
#       main="Occupancy", 
#       xlab="Theoretical Quantiles", 
#       ylab="Standardised Residuals")
#qqline(occuWright$res) #add a straight diagonal line to the plot

# detection residuals
#qqnorm(detWright$res, 
#       main="Occupancy", 
#       xlab="Theoretical Quantiles", 
#       ylab="Standardised Residuals")
#qqline(detResidW$res)
```
      
      
      
      
      
      
      
      
      
      
      
      





# Species-wise summary

```{r plot-occu-detProb-setup, message=FALSE, warning=FALSE, echo=FALSE}

### Before plotting, we format all models such that the species-wise plot code is short and sweet


require(ggplot2)
require(labelmachine)
require(sparta)
require(dplyr)
require(plyr)
require(boot)
#require(reshape2)

### Source plotting function
source("./fig-plot.Occdet_G.R")

# Change ModelNames vetor to species names with spaces
spp_names <- gsub(ModelNames, patt = "_", repl=" ")
spp_names <- sub(spp_names, patt = "c.album", repl="c-album")

# Name the lists of outputs their species names
names(mixLL_list)[1:12] <- spp_names
names(mixLL2_list)[1:12] <- spp_names


# Create a dictionary translating Latin names to common names
dict <- new_lama_dictionary(common_names = 
                              c("Aglais urticae" = 'Small Tortoiseshell',
                                "Carterocephalus palaemon" = 'Chequered Skipper',
                                "Erebia aethiops" = 'Scotch Argus',
                                "Erebia epiphron" = 'Mountain Ringlet',
                                "Limenitis camilla" = 'White Admiral',
                                "Lycaena phlaeas" = 'Small Copper',
                                "Melitaea athalia" = 'Heath Fritillary',
                                "Pararge aegeria" = 'Speckled Wood',
                                "Pieris rapae" = 'Small White',
                                "Polygonia c-album" = 'Comma',
                                "Polyommatus bellargus" = 'Adonis Blue',
                                "Thecla betulae" = 'Brown Hairstreak'))

source("./fig-plot.Occdet_G.R")
```


```{r plot-detprob-outputs-setup, message=FALSE, warning=FALSE, eval=FALSE, echo=FALSE}
require(ggplot2)
require(labelmachine)
require(sparta)
require(dplyr)
require(plyr)
require(boot)


# Change ModelNames vetor to species names with spaces
spp_names <- gsub(ModelNames, patt = "_", repl=" ")
spp_names <- sub(spp_names, patt = "c.album", repl="c-album")

# Name the lists of outputs their species names
names(mixLL_list)[1:12] <- spp_names
names(mixLL2_list)[1:12] <- spp_names


# Create a dictionary translating Latin names to common names
dict <- new_lama_dictionary(common_names = 
                              c("Aglais urticae" = 'Small Tortoiseshell',
                                "Carterocephalus palaemon" = 'Chequered Skipper',
                                "Erebia aethiops" = 'Scotch Argus',
                                "Erebia epiphron" = 'Mountain Ringlet',
                                "Limenitis camilla" = 'White Admiral',
                                "Lycaena phlaeas" = 'Small Copper',
                                "Melitaea athalia" = 'Heath Fritillary',
                                "Pararge aegeria" = 'Speckled Wood',
                                "Pieris rapae" = 'Small White',
                                "Polygonia c-album" = 'Comma',
                                "Polyommatus bellargus" = 'Adonis Blue',
                                "Thecla betulae" = 'Brown Hairstreak'))

```

## *Aglais urticae* (Small Tortoiseshell)

### Occupancy
```{r plot-occu-outputs-A-urticae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[1]], mixLL2_list[[1]], catLL_list[[1]]) +
  ggtitle(label = "Aglais urticae", 
          subtitle = paste(lama_translate_("Aglais urticae", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  theme(text=element_text(size=15)) 
```

     
        
        
        
        

*Aglais urticae* (Small Tortoiseshell) is one of the UKs most widespread butterflies, occurring throughout the British Isles, which is in line with the latter part of the time series (post 1976; but note it has declined by approx 15% since). Nonetheless, there is no anecdotal evidence for a 100% increase between the 1960s and 1980s indicating that the occupancy is underestimated in the early part of the time series.   
   
   
   
**In other words it appears neither of the model formulations are accurately explaining early trends for this species**


   

### Detection probability
```{r plot-detprob-outputs-A-urticae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}
plot_DetectionOverTime(catLL_list[[1]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  #ggtitle(label = (expression(paste(italic("Aglais urticae"), " - Model A")))) #, 
  #        subtitle = paste(lama_translate_("Aglais urticae", dict, "common_names"))) + 
  #theme(plot.title=element_text(face="italic")) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[1]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[1]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))
```


      
      
      
      
BNM and UKBMS detection probability trends are similar for all three models after 1976. However, NHCs detection probability estimates differ between models such that it is the data type with the highest detection probability (unlikely) in Model A, and the lowest in model C; still the detection probability estimates are around 0.07 for model C. If we assume that all NHC visits produced single species lists (which the majority did), this would mean that every 5-10% of museum visits collected a small tortoiseshell, which again seems unlikely as that in turn would mean that approx. the same percentage of the entire collection would be small tortoiseshells).



### Visits with records

```{r plot-siteprop-A-urticae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Aglais urticae*")

test$Model <- as.factor(test$Model)


require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"),
                  alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF",
                                    "#20A387FF",
                                    "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF",
                                    "#20A387FF",
                                    "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") +
  xlab("Year") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
      #theme(legend.position="bottom") +
  xlab("Year") + 
  ylab("Proportion of visited sites with records") + 
  ggtitle("Aglais urticae") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-A-urticae, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4}
# Load required packages
require(dplyr)
require(plyr)
require(ggplot2)
require(viridis)
require(readr)
require(ggtext)

# Load spp_vis data (any would do), which contain all visits and whether or not each of the three species have been observed at that visit
spp_vis <- read.csv('../data/formattedData/spp_vis_merged_2020-04-07.csv', header=T, na.strings=c("","NA"))

# Load occDetdata giving each visits' site ID (grid cell), list length (numper of species observed) and time period (in this case year)
occDetdata_merged_2020_04_07 <- read.csv('../data/formattedData/occDetdata_merged_2020-04-07.csv', header=T, na.strings=c("","NA"))

# Merge by visit
spp_vis <- merge(spp_vis, occDetdata_merged_2020_04_07, by="visit")

# Find unique visits
spp_vis <- unique(spp_vis)

# Create a species column
spp_vis$species <- as.factor(c("NA"))

### Make L column a source column
# Change all List lengths between 0 and 22222 to 1 (all these are BNM data)
spp_vis$L <- replace(spp_vis$L, spp_vis$L > 0 & spp_vis$L < 22222, 1)
# Make L a factor
spp_vis$L <- as.factor(spp_vis$L)
# Rename factor levels according to data type
spp_vis$L <- revalue(spp_vis$L, c("1"="BNM", "22222"="UKBMS", "33333"="NHCs"))
# Rename column to Source
colnames(spp_vis)[colnames(spp_vis)=="L"] <- "Source"


# Correct time period so it starts at 1900
spp_vis$year <- spp_vis$TP+1899



########################################################################

# Subset each species' records into dataframes and specify species identity
spp_vis_A_urticae <- subset(spp_vis, Aglais.urticae == "TRUE")
spp_vis_A_urticae$species <- c("italic('A. urticae')")



 ggplot(spp_vis_A_urticae, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Aglais urticae") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))


```





### DIC & Bayesian *p*
```{r DIC-A-urticae, cache=TRUE, message=FALSE, echo=FALSE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[1]]$BUGSoutput$DIC), 
                                (mixLL_list[[1]]$BUGSoutput$DIC),
                                (mixLL2_list[[1]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[1, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[1, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[1, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[1, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[1, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[1, 7]

knitr::kable(summary_out)
```






* Bayesian p:     
    + very poor for variance      
    + OK for mean    
    + Model A (Cat list length) best    
* DIC: Model C (mic list length 2) best   







## *Carterocephalus palaemon* (Chequered skipper)

### Occupancy
```{r plot-occu-outputs-C-palaemon, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[2]], mixLL2_list[[2]], catLL_list[[2]]) +
  ggtitle(label = "Carterocephalus palaemon", 
          subtitle = paste(lama_translate_("Carterocephalus palaemon", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15))  
```
   
   
   
   
The Chequered Skipper is confined to north-west Scotland where it was first discovered in 1939 (distribution is centered on Fort William). It formerly occurred in central and east England, but started to decline in "the early  part of the 20th century" (Millennium atlas) with East Midlands as the last English stronghold where it was "fairly abundant until the 1950s". Thereafter remaining English populations suffered rapid declines and it was declared extinct in England in 1976. 
   
   
**Occupancy estimates are in line with anecdotal evidence for historic distribution changes and very similar across models**   
    
   
### Detection probability
```{r plot-detprob-outputs-C-palaemon, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}
plot_DetectionOverTime(catLL_list[[2]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[2]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[2]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```
   
   
   
   
Detection probabilities when compared between models are similar to previous species. However, I would expect detection probabilities to be higher? Although it is quite hard to spot (skipper = rapid flier etc.). BUT I would expect the detection probability for NHCs in Model C to be much higher than it is.... 
   
   
**General thought: detection probability does not appear to influence occupancy estimates that much?**   
    
   



### Visits with records
```{r plot-siteprop-C-palaemon, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Carterocephalus palaemon*")

test$Model <- as.factor(test$Model)


require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"),
                  alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") +
  xlab("Year") +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
      #theme(legend.position="bottom") +
  xlab("Year") + 
  ylab("Proportion of visited sites with records") + 
  ggtitle("Carterocephalus palaemon") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()

```


### Record numbers
```{r plot-record-numbers-C-palaemon, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_C_palaemon <- subset(spp_vis, Carterocephalus.palaemon == "TRUE")
spp_vis_C_palaemon$species <- c("italic('C. palaemon')")

 ggplot(spp_vis_C_palaemon, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Carterocephalus palaemon") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
 
```


### DIC & Bayesian *p*
```{r DIC-C-palaemon, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[2]]$BUGSoutput$DIC), 
                                (mixLL_list[[2]]$BUGSoutput$DIC),
                                (mixLL2_list[[2]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[2, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[2, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[2, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[2, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[2, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[2, 7]

knitr::kable(summary_out)
```


* Bayesian p:     
    + Poor for variance: A best      
    + OK for mean: C best          
* DIC: Model B (mic list length) best   




## *Erebia aethiops* (Scotch Argus)

### Occupancy
```{r plot-occu-outputs-E-aethiops, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[3]], mixLL2_list[[3]], catLL_list[[3]]) +
  ggtitle(label = "Erebia aethiops", 
          subtitle = paste(lama_translate_("Erebia aethiops", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.25), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15))  
```


### Detection probability
```{r plot-detprob-outputs-E-aethiops, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}
plot_DetectionOverTime(catLL_list[[3]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[3]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[3]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-E-aethiops, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Erebia aethiops*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Erebia aethiops") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-E-aethiops, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_E_aethiops <- subset(spp_vis, Erebia.aethiops == "TRUE")
spp_vis_E_aethiops$species <- c("italic('E. aethiops')")

 ggplot(spp_vis_E_aethiops, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Erebia aethiops") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-E-aethiops, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[3]]$BUGSoutput$DIC), 
                                (mixLL_list[[3]]$BUGSoutput$DIC),
                                (mixLL2_list[[3]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[3, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[3, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[3, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[3, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[3, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[3, 7]

knitr::kable(summary_out)
```




## *Erebia epiphron* (Mountain Ringlet)

### Occupancy
```{r plot-occu-outputs-E-epiphron, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[4]], mixLL2_list[[4]], catLL_list[[4]]) +
  ggtitle(label = "Erebia epiphron", 
          subtitle = paste(lama_translate_("Erebia epiphron", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-E-epiphron, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}
plot_DetectionOverTime(catLL_list[[4]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[4]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[4]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-E-epiphron, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Erebia epiphron*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Erebia epiphron") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-E-epiphron, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_E_epiphron <- subset(spp_vis, Erebia.epiphron == "TRUE")
spp_vis_E_epiphron$species <- c("italic('E. epiphron')")

 ggplot(spp_vis_E_epiphron, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Erebia epiphron") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-E-epiphron, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[4]]$BUGSoutput$DIC), 
                                (mixLL_list[[4]]$BUGSoutput$DIC),
                                (mixLL2_list[[4]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[4, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[4, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[4, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[4, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[4, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[4, 7]

knitr::kable(summary_out)
```















## *Limenitis camilla* (White Admiral)

### Occupancy
```{r plot-occu-outputs-L-camilla, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[5]], mixLL2_list[[5]], catLL_list[[5]]) +
  ggtitle(label = "Limenitis camilla", 
          subtitle = paste(lama_translate_("Limenitis camilla", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.4), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-L-camilla, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}
plot_DetectionOverTime(catLL_list[[5]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[5]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[5]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))
```



### Visits with records
```{r plot-siteprop-L-camilla, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Limenitis camilla*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.25), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Limenitis camilla") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-L-camilla, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_L_camilla <- subset(spp_vis, Limenitis.camilla == "TRUE")
spp_vis_L_camilla$species <- c("italic('L. camilla')")

 ggplot(spp_vis_L_camilla, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Limenitis camilla") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-L-camilla, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[5]]$BUGSoutput$DIC), 
                                (mixLL_list[[5]]$BUGSoutput$DIC),
                                (mixLL2_list[[5]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[5, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[5, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[5, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[5, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[5, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[5, 7]

knitr::kable(summary_out)
```















## *Lycaena phlaeas* (Small copper)

### Occupancy
```{r plot-occu-outputs-L-phlaeas, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[6]], mixLL2_list[[6]], catLL_list[[6]]) +
  ggtitle(label = "Lycaena phlaeas", 
          subtitle = paste(lama_translate_("Lycaena phlaeas", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-L-phlaeas, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}
plot_DetectionOverTime(catLL_list[[6]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[6]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[6]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))
```



### Visits with records
```{r plot-siteprop-L-phlaeas, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Lycaena phlaeas*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.8), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Lycaena phlaeas") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-L-phlaeas, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_L_phlaeas <- subset(spp_vis, Lycaena.phlaeas == "TRUE")
spp_vis_L_phlaeas$species <- c("italic('L. phlaeas')")

 ggplot(spp_vis_L_phlaeas, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Lycaena phlaeas") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-L-phlaeas, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[6]]$BUGSoutput$DIC), 
                                (mixLL_list[[6]]$BUGSoutput$DIC),
                                (mixLL2_list[[6]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[6, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[6, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[6, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[6, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[6, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[6, 7]

knitr::kable(summary_out)
```















## *Melitaea athalia* (Heath Fritillary)

### Occupancy
```{r plot-occu-outputs-M-athalia, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[7]], mixLL2_list[[7]], catLL_list[[7]]) +
  ggtitle(label = "Melitaea athalia", 
          subtitle = paste(lama_translate_("Melitaea athalia", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-M-athalia, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}

plot_DetectionOverTime(catLL_list[[7]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[7]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[7]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-M-athalia, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Melitaea athalia*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Melitaea athalia") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-M-athalia, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_M_athalia <- subset(spp_vis, Melitaea.athalia == "TRUE")
spp_vis_M_athalia$species <- c("italic('M. athalia')")

 ggplot(spp_vis_M_athalia, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Melitaea athalia") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-M-athalia, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[7]]$BUGSoutput$DIC), 
                                (mixLL_list[[7]]$BUGSoutput$DIC),
                                (mixLL2_list[[7]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[7, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[7, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[7, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[7, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[7, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[7, 7]

knitr::kable(summary_out)
```














4. *Polyommatus bellargus*   
5. *Thecla betulae*   
7. *Polygonia c-album*   
10. *Pararge aegeria*   
12. *Pieris rapae*   


## *Pararge aegeria* (Speckled Wood)

### Occupancy
```{r plot-occu-outputs-P-aegeria, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[8]], mixLL2_list[[8]], catLL_list[[8]]) +
  ggtitle(label = "Pararge aegeria", 
          subtitle = paste(lama_translate_("Pararge aegeria", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-P-aegeria, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}

plot_DetectionOverTime(catLL_list[[8]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[8]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[8]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-P-aegeria, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Pararge aegeria*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.8), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Pararge aegeria") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-P-aegeria, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_P_aegeria <- subset(spp_vis, Pararge.aegeria == "TRUE")
spp_vis_P_aegeria$species <- c("italic('P. aegeria')")

 ggplot(spp_vis_P_aegeria, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Pararge aegeria") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-P-aegeria, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[8]]$BUGSoutput$DIC), 
                                (mixLL_list[[8]]$BUGSoutput$DIC),
                                (mixLL2_list[[8]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[8, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[8, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[8, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[8, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[8, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[8, 7]

knitr::kable(summary_out)
```












   



## *Pieris rapae* (Small/Cabbage White)

### Occupancy
```{r plot-occu-outputs-P-rapae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[9]], mixLL2_list[[9]], catLL_list[[9]]) +
  ggtitle(label = "Pieris rapae", 
          subtitle = paste(lama_translate_("Pieris rapae", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-P-rapae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}

plot_DetectionOverTime(catLL_list[[9]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[9]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[9]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-P-rapae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Pieris rapae*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.8), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Pieris rapae") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-P-rapae, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_P_rapae <- subset(spp_vis, Pieris.rapae == "TRUE")
spp_vis_P_rapae$species <- c("italic('P. rapae')")

 ggplot(spp_vis_P_rapae, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Pieris rapae") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-P-rapae, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[9]]$BUGSoutput$DIC), 
                                (mixLL_list[[9]]$BUGSoutput$DIC),
                                (mixLL2_list[[9]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[9, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[9, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[9, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[9, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[9, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[9, 7]

knitr::kable(summary_out)
```




















## *Polygonia c-album* (Comma)

### Occupancy
```{r plot-occu-outputs-Polygonia-c-album, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[10]], mixLL2_list[[10]], catLL_list[[10]]) +
  ggtitle(label = "Polygonia c-album", 
          subtitle = paste(lama_translate_("Polygonia c-album", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-P-c-album, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}

plot_DetectionOverTime(catLL_list[[10]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[10]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[10]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-P-c-album, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Polygonia c-album*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.7), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Polygonia c-album") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-P-c-album, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_P_c_album <- subset(spp_vis, Polygonia.c.album == "TRUE")
spp_vis_P_c_album$species <- c("italic('P. c-album')")

 ggplot(spp_vis_P_c_album, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Polygonia c-album") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-P-c-album, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[10]]$BUGSoutput$DIC), 
                                (mixLL_list[[10]]$BUGSoutput$DIC),
                                (mixLL2_list[[10]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[10, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[10, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[10, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[10, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[10, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[10, 7]

knitr::kable(summary_out)
```













4. *Polyommatus bellargus*   
5. *Thecla betulae*   






## *Polyommatus bellargus* (Adonis Blue)

### Occupancy
```{r plot-occu-outputs-P-bellargus, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[11]], mixLL2_list[[11]], catLL_list[[11]]) +
  ggtitle(label = "Polyommatus bellargus", 
          subtitle = paste(lama_translate_("Polyommatus bellargus", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-P-bellargus, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}

plot_DetectionOverTime(catLL_list[[11]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[11]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[11]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-P-bellargus, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Polyommatus bellargus*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.2), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Polyommatus bellargus") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-P-bellargus, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_P_bellargus <- subset(spp_vis, Polyommatus.bellargus == "TRUE")
spp_vis_P_bellargus$species <- c("italic('P. bellargus')")

 ggplot(spp_vis_P_bellargus, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Polyommatus bellargus") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-P-bellargus, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[11]]$BUGSoutput$DIC), 
                                (mixLL_list[[11]]$BUGSoutput$DIC),
                                (mixLL2_list[[11]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[11, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[11, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[11, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[11, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[11, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[11, 7]

knitr::kable(summary_out)
```





 




## *Thecla betulae* (Brown Hairstreak)

### Occupancy
```{r plot-occu-outputs-P-betulae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
### Source plotting function
source("./fig-plot.Occdet_G.R")

plot.occDet_G(mixLL_list[[12]], mixLL2_list[[12]], catLL_list[[12]]) +
  ggtitle(label = "Thecla betulae", 
          subtitle = paste(lama_translate_("Thecla betulae", dict, "common_names"))) + 
  theme(plot.title=element_text(face="italic")) +
  scale_y_continuous(limits=c(0, 0.5), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme(text=element_text(size=15)) 
```



### Detection probability
```{r plot-detprob-outputs-P-betulae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=3}

plot_DetectionOverTime(catLL_list[[12]],
                              min.yr = 1900, 
                                    legend_labels =
                         c("    BNM/Year_Effect  ", 
                                                    "UKBMS", 
                                                   "NHCs"),
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model A") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12)) 

plot_DetectionOverTime(mixLL_list[[12]], 
                       min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model B") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  theme(text=element_text(size=12))

plot_DetectionOverTime(mixLL2_list[[12]],
                              min.yr = 1900, 
                       legend_labels = c("BNM_LL1/Year effect", 
                                         "UKBMS", 
                                         "NHCs", 
                                         "BNM_LL5"), 
                       mixLL2 = TRUE,                  ### Important: specify its mixLL2
                       legend_title = "Data Type") +
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  ggtitle("Model C") +
  theme(text=element_text(size=12))

```



### Visits with records
```{r plot-siteprop-P-betulae, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4}
sppMeanSitesYear$Species <- as.factor(sppMeanSitesYear$Species)
test <- subset(sppMeanSitesYear, Species == "*Thecla betulae*")

test$Model <- as.factor(test$Model)

require("ggtext")
ggplot(data = test,) +
  geom_ribbon(aes_string(fill = "Model", 
                    x = "Year", 
                    y = "sim_MeanSitesPerYear", 
                             ymin = "sim_lower", ymax = "sim_upper"), alpha = 0.3) +
  scale_fill_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) + 
  geom_line(aes(x = Year, y = sim_MeanSitesPerYear, colour = Model), size=0.7) + 
          scale_colour_manual(values =c("#440154FF", "#20A387FF", "#E69F00")) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), linetype = "dotted", colour=c("black"), size = 1.5) +
  geom_line(aes(x=Year, y=obs_meanSitesPerYear), colour=c("black"), size = 0.4) +
  geom_vline(xintercept = 1976, linetype = "dashed") +   # Vertical line = 1976
  ylab("Proportion of visited sites with records") + xlab("Year") +
  scale_y_continuous(limits=c(0, 0.1), expand = c(0, 0)) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Thecla betulae") + 
  theme(plot.title=element_text(face="italic")) +
  theme_bw()
```


### Record numbers
```{r plot-record-numbers-P-betulae, message=FALSE, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
# Subset each species' records into dataframes and specify species identity
spp_vis_T_betulae <- subset(spp_vis, Thecla.betulae == "TRUE")
spp_vis_T_betulae$species <- c("italic('T. betulae')")

 ggplot(spp_vis_T_betulae, aes(x=year, fill=Source, color=Source)) + 
  scale_fill_manual(values = alpha(c("#453781FF", 
                                     "#20A386FF", 
                                     "#DCE318FF"), 0.5)) +
  scale_color_manual(values = alpha(c("#453781FF", 
                                      "#20A386FF", 
                                      "#DCE318FF"), 0.5)) +
  geom_bar(position="identity", alpha=0.7) + 
  labs(y="Records", x = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  # sqrt y scale and expand scale to no space below, 5% above
  scale_y_sqrt(expand = expansion(mult = c(0, .05))) +
  scale_x_continuous(breaks=seq(1900, 2020, 20), 
                     limits = c(1900, 2025), 
                     expand = c(0, 0)) +
  ggtitle("Thecla betulae") + 
  theme(plot.title=element_text(face="italic")) +
  theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))
```


### DIC & Bayesian *p*
```{r DIC-P-betulae, cache=TRUE, message=FALSE, echo=FALSE, cache=TRUE}
# Create empty data frame to populate
summary_out <- data.frame(Model = as.factor(c("catLL", "mixLL", "mixLL2")),
                          DIC = as.numeric(NA),
                          Bayesian_p = as.numeric(NA),
                          Bayesian_p_var = as.numeric(NA)) 


  summary_out[1:3, "DIC"] <-  c((catLL_list[[12]]$BUGSoutput$DIC), 
                                (mixLL_list[[12]]$BUGSoutput$DIC),
                                (mixLL2_list[[12]]$BUGSoutput$DIC))  
  summary_out[1, "Bayesian_p"] <-  ModelBp_Summ[12, 2]
  summary_out[2, "Bayesian_p"] <-  ModelBp_Summ[12, 3]
  summary_out[3, "Bayesian_p"] <-  ModelBp_Summ[12, 4]
  summary_out[1, "Bayesian_p_var"] <-  ModelBp_Summ[12, 5]
  summary_out[2, "Bayesian_p_var"] <-  ModelBp_Summ[12, 6]
  summary_out[3, "Bayesian_p_var"] <-  ModelBp_Summ[12, 7]

knitr::kable(summary_out)
```
















###### Shiny plots - not used but to annotate
```{r SHINY-plot-occu-outputs, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
require(shiny)
require(ggplot2)
require(labelmachine)
require(sparta)
require(dplyr)

### Source plotting function
source("./fig-plot.Occdet_G.R")

# Change ModelNames vetor to species names with spaces
spp_names <- gsub(ModelNames, patt = "_", repl=" ")
spp_names <- sub(spp_names, patt = "c.album", repl="c-album")

# Name the lists of outputs their species names
names(mixLL_list)[1:12] <- spp_names
names(mixLL2_list)[1:12] <- spp_names


# Create a dictionary translating Latin names to common names
dict <- new_lama_dictionary(common_names = 
                              c("Aglais urticae" = 'Small Tortoiseshell',
                                "Carterocephalus palaemon" = 'Chequered Skipper',
                                "Erebia aethiops" = 'Scotch Argus',
                                "Erebia epiphron" = 'Mountain Ringlet',
                                "Limenitis camilla" = 'White Admiral',
                                "Lycaena phlaeas" = 'Small Copper',
                                "Melitaea athalia" = 'Heath Fritillary',
                                "Pararge aegeria" = 'Speckled Wood',
                                "Pieris rapae" = 'Small White',
                                "Polygonia c-album" = 'Comma',
                                "Polyommatus bellargus" = 'Adonis Blue',
                                "Thecla betulae" = 'Brown Hairstreak'))


# Plot occupancy estimates as a shiny app (currently two models per species)

shinyApp(
  ui = fluidPage(
    # Select input
    selectInput("species_choice", # call the variable storing values (spp names) "species_choice"
                label = "Select species", # header text of dropdown list
                choices = spp_names, # vector of all values in dropdown list—in (species)
                selected ="Aglais urticae"), # dropdown defaults to Aglais urticae
    plotOutput("occuPlot")
  ),
  server = function(input, output) {
    output$occuPlot <- renderPlot({
      
      vec <- paste(input$species_choice)  
      vec_labeled <- lama_translate_(vec, dict, "common_names")  
      
      plot.occDet_G(mixLL_list[[input$species_choice]], mixLL2_list[[input$species_choice]]) +
        ggtitle(label = paste(input$species_choice), 
                subtitle = paste(vec_labeled)) + 
        theme(plot.title=element_text(face="italic")) +
        theme(text=element_text(size=18))
    })
  },
  options = list(height = 500)
)
```


 
```{r SHINY-plot-det-outputs, message=FALSE, echo=FALSE, fig.show='hold', out.width='50%', eval=FALSE}

require(shiny)
require(ggplot2)
require(sparta)
require(dplyr)

 
# Plot detection probability estimates as a shiny app (currently two models per species)
 
selectInput("species_choice", 
            label = "Select species",
            choices = spp_names, 
            selected ="Aglais urticae")
fluidPage(fluidRow(
#  column(5, 
#         renderPlot(
#           plot_DetectionOverTime(catLL_list[[input$species_choice]], 
#                                  min.yr = 1900) 
#             theme(legend.position = "none") +
#             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
#             scale_x_continuous(expand = c(0, 0)) +
#             ggtitle(label = paste(input$species_choice), 
#                subtitle = "Model A") + 
#             theme(plot.title=element_text(face="italic")) +
#             theme(text=element_text(size=16))
#         )
#         ),
  column(5, 
         renderPlot(
           plot_DetectionOverTime(mixLL_list[[input$species_choice]], 
                                  min.yr = 1900) + 
             theme(legend.position = "none") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
             ggtitle(label = paste(input$species_choice), 
                     subtitle = "Model B") + 
             theme(plot.title=element_text(face="italic")) +
             theme(text=element_text(size=16))
         )
         ),
  column(7, 
         renderPlot(
           plot_DetectionOverTime(mixLL2_list[[input$species_choice]], 
                                  min.yr = 1900, 
                                  legend_labels = c("BNM_LL1/Year_Effect", 
                                                    "UKBMS", 
                                                    "NHCs", 
                                                    "BNM_LL5"), 
                                  legend_title = "Data Type") +
             scale_y_continuous(limits=c(0, 1), expand = c(0, 0)) +
             scale_x_continuous(expand = c(0, 0)) +
             ggtitle(label = paste(input$species_choice, 
                subtitle = "Model C") + 
             theme(axis.title.y = element_blank(),
                   axis.text.y = element_blank(),
                   text=element_text(size=16))
           )
         ) 
  )
)
 
```